### 2022-11-07 Code/Exploratory_Analysis_CAP_DAP.Rmd
Committing all the data and the start of the code

---
title: "Explorative Data Analysis Climate Perceptions in Countries"
author: "Rae Dunbar"
date: "11/7/2022"
output: html_document
---
For this project theere are several csv's that came from three excel sheets. The first one is emissions per country from 1990-2019. The second is a survey about countries climate perceptions that Yale used for a research paper. And the third is from a Facebook survey of several countries perceptions of climate change. The facebook survey data put its data in wide format using several sheets, so I had to make each sheet into a separate csv. That dataset will be the most burdensome to clean up.

Lets start with thee emissions data.

## Setting R up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

## Starting with the Emissions Data

### Pulling the Data in

```{r}
emissions<-read.csv(here("Data for the Code", "CAIT_percapita_allcountries.csv"), stringsAsFactors = TRUE)
```

### viewing the Data

What are the headers?
```{r}
names(emissions)
```
I appears the first few columns after the country name and unit are the dates, starting with 1990 and going until 2019. There are alot of other columns after that however, that are unclear of what data they hold.

```{r}

```

### 2022-11-08 Code/Exploratory_Analysis_CAP_DAP.Rmd
uplodaing metadata, altering code making progress

@@
 
-Lets start with thee emissions data.
+Lets start with the emissions data.
 
@@
 
+Lets learn more.
+
 ```{r}
+str(emissions)
+```
+Another issue I have noticed is all the data is coming in as factor data. We can check this with glimpse()
 
+```{r}
+glimpse(emissions)
 ```
+Here we can see both the odd amount of empty data columns and that they are all listed as factor data, which we will need to change to numerical data.
+
+When looking at the full data set, we can also see that there are two rows at the bottom that are unecessary.They appear to be metadata.
+
+##Fixing the Data
+
+First we only want columns with data, so we need to select the rows we desire.
+
+### Emitting the Empty Columns
+```{r}
+emissions<- emissions %>% select("Country.Region", "X1990":"X2019")
+emissions
+
+#The : actually worked inside select, which is neat.
+```
+Now we should rename the column names.
+
+### Renaming the Column Names
+
+THIS DIDNT WORK WAAAAA?
+```{r}
+rename(emissions, country=Country.Region)  
+# rename allows one to rename individual column names and does not require the coder to type all of the columns.
+```
+In a perfect world I would rename all of the dates to 1990 and so forth, rather than X1990.The issue is, R thinks when I type 1990 I am referring to a number, so the current column names will have to stay.
+
+When looking at the full data set, we can also see that there are two rows at the bottom that are unecessary.They appear to be metadata. We need to remove these from the dataframe.
+
+## Emitting the Last Two Rows
+```{r}
+emissions <- emissions %>% slice(-194,-195)
+# Gets rid of the last two rows.
+```
+
+Lastly, we need to make all the date columns into
+
+### Changing Factors to Numeric
+
+This code below messes the number ups somehow.
+```{r}
+#emissions$X1990<- as.numeric(emissions$X1990)
+```
+NA's were introduced by corecion, because some of the data contains the word "FALSE" in it.
+Which, is just changed to an NA.
+I believe this is the best course of action anyway, since the metadata does not explain what FALSE means.
+
+To not have to run the above code 30 times, we can try a loop, though it did not work. 
+```{r}
+#for(i in emissions$X1990:emissions$X2019){
+#i <- as.numeric(as.character(i))
+#}
+```
+
+Lets try lapply and slapply instead.
+If we are to use them, we need the first column to not be a factor temporarily, soo it does not get turned into numeric data, which woould be bad, given they are country names.
+
+This will work but I had to do a weird thing.
+```{r}
+#So, instead I separated the data frames into two, just changed the data types of the data frame with the dates in them, and the used cbind to glue them back together into one dataframe! 
+emissions1<- emissions %>% select(Country.Region)
+emissionsfix<- emissions %>% select(-Country.Region)
+
+indx <- sapply(emissionsfix, is.factor)
+emissionsfix[indx] <- lapply(emissionsfix[indx], function(x) as.numeric(as.character(x)))
+
+emissions<- cbind(emissions1, emissionsfix)
+```
+
+Now the data is all swell and we can go to checking for outliers and running tests.
+
+## Graphing the Data
+We want to compare the emissions of Denmark to the United States from 1990 to 2019, so we are going to want a scatter plot.
+
+Now we have the added issue that this is actually wide format not long.
+```{r}
+emissions_graph<- emissions %>% filter(Country.Region=="Denmark" | Country.Region=="United States")
+ggplot(emissions, aes(x=))
+```
+
+## Pivoting to long format
+
+This worked!
+```{r}
+emissions_graph<-emissions_graph %>% pivot_longer(cols=c('X1990':'X2019'),
+                    names_to='year',
+                    values_to='emissions')
+```
+
+### Graphing
+
+```{r}
+ggplot(emissions_graph, aes(x=year, y=emissions, colour=Country.Region)) + geom_point()+theme_bw()
+```
+
+It looks like Denmark has significantly less emissions than the U.S. There are some odd spikes in Denmarks emissions, but nothing that seems to be an outlier.
+
+## Running the Test
+
+We have our explanatory which is categorical, and our response of emissions, which is continuous. So, we want a two sample t test.
+```{r}
+t.test(emissions~Country.Region, data=emissions_graph)
+```
+The p value is very small, so the difference is significant. The United States released significantly more greenhouse gases than Denmark.
+
+## Rate

### 2022-11-08 Code/Exploratory_Analysis_CAP_DAP.Rmd
Committing Rmd with all the emissions code, plus the data for Facebook

--- 2022-11-07 Code/Exploratory_Analysis_CAP_DAP.Rmd
@@
 ---
-For this project theere are several csv's that came from three excel sheets.
+For this project there are several csv's that came from three excel sheets.
 The first one is emissions per country from 1990-2019.
@@
 And the third is from a Facebook survey of several countries perceptions of climate change.
-The facebook survey data put its data in wide format using several sheets, so I had to make each sheet into a separate csv.
-That dataset will be the most burdensome to clean up.
+The facebook survey data put its data in wide format using several sheets, so there are several separate csvs.
+That data set will be the most burdensome to clean up.
 
-Lets start with thee emissions data.
+Lets start with the emissions data.
 
@@
 ```
-I appears the first few columns after the country name and unit are the dates, starting with 1990 and going until 2019.
-There are alot of other columns after that however, that are unclear of what data they hold.
+It appears the first few columns after the country name and unit are the dates, starting with 1990 and going until 2019.
+There are a lot of other columns after that however, that are unclear of what data they hold.
+I will have to delete those and I will also need to make this into long format.
+
+Lets learn more.
+
+```{r}
+str(emissions)
+```
+Another issue I have noticed is all the data is coming in as factor data. We can check this with glimpse()
+
+```{r}
+glimpse(emissions)
+```
+Here we can see both the odd amount of empty data columns and that they are all listed as factor data, which we will need to change to numerical data.
+If we change it to numerical data we will also need to get rid of the X in front of the dates.
+
+When looking at the full data set, we can also see that there are two rows at the bottom that are unnecessary.They appear to be metadata.
+
+What need to be fixed:
+1. removing the blank columns
+2. renaming the column names
+3. removing the last two rows
+4. turning the data into long format
+5. getting rid of the X's at the beginning of the dates
+
+##Fixing the Data
+
+First we only want columns with data, so we need to select the rows we desire.
+
+### Emitting the Empty Columns
+```{r}
+emissions<- emissions %>% select("Country.Region", "X1990":"X2019")
+emissions
+
+#The : actually worked inside select, which is neat.
+```
+Now we should rename the column names.
+
+### Renaming the Column Names
+
+```{r}
+emissions<-rename(emissions, country=Country.Region)  
+# rename allows one to rename individual column names and does not require the coder to type all of the columns.
+#Make sure you assign it to your data frame!!!!!!
+```
+In a perfect world I would rename all of the dates to 1990 and so forth, rather than X1990.The issue is, R thinks when I type 1990 I am referring to a number, so the current column names will have to stay.
+We also need it in long format anyway, so just fixing the Country.Region one is good.
+
+When looking at the full data set, we can also see that there are two rows at the bottom that are unnecessary.They appear to be metadata. We need to remove these from the dataframe.
+
+## Emitting the Last Two Rows
+```{r}
+emissions <- emissions %>% slice(-194,-195)
+# Gets rid of the last two rows.
+```
+
+Lastly, we need to make all the date columns into numeric rather than factor data, so when we put it in long format R knows they are numbers.
+
+### Changing Factors to Numeric
+
+This code below messes the number ups somehow.
+```{r}
+#emissions$X1990<- as.numeric(emissions$X1990)
+```
+NA's were introduced by corecion, because some of the data contains the word "FALSE" in it.
+Which, is just changed to an NA.
+I believe this is the best course of action anyway, since the metadata does not explain what FALSE means.
+
+To not have to run the above code 30 times, we can try a loop, though it did not work. 
+```{r}
+#for(i in emissions$X1990:emissions$X2019){
+#i <- as.numeric(as.character(i))
+#}
+```
+
+Lets try lapply and slapply instead.
+If we are to use them, we need the first column to not be a factor temporarily, soo it does not get turned into numeric data, which would be bad, given they are country names.
+
+```{r}
+#So, instead I separated the data frames into two, just changed the data types of the data frame with the dates in them, and then used cbind to glue them back together into one dataframe! 
+emissions1<- emissions %>% select(country)
+emissionsfix<- emissions %>% select(-country)
+indx <- sapply(emissionsfix, is.factor)
+emissionsfix[indx] <- lapply(emissionsfix[indx], function(x) as.numeric(as.character(x)))
+
+#sapply (the data frame, what you want to do to it).
+So you assign indx to all the columns that are factors in emissionsfix.
+And then lapply does what you want it to do to every single column.
+So you say I want you to work with emissionsfix[at the index indx] and then do this function, which is to make all of it to character and then to a number.
+You cannot go from factor to number.
+
+emissions<- cbind(emissions1, emissionsfix)
+```
+NA's were introduced by coercion into the countries that have FALSE listed as their emissions, because their emissions could not be recorded.
+
+Now we need to chagne the data from wide to long format.
+
+## Pivoting to long format
+
+This worked!
+```{r}
+emissions_long<-emissions %>% pivot_longer(cols=c('X1990':'X2019'),
+                    names_to='year',
+                    values_to='emissions')
+```
+
+Now we want to get rid of the X infront of the X1990 etc.
+
+## Getting rid of the X
+
+For the technique that I came up with it is easier with fewer countries. So, we want to compare the emissions of Denmark to the United States from 1990 to 2019.
+
+First we need to filter the data so we can just look at the two states of interest. 
+```{r}
+emissions_US_Denmark<- emissions_long %>% filter(country=="Denmark" | country =="United States")
+```
+
+```{r}
+
+emissions_US_Denmark$year<- 'NA'
+emissions_US_Denmark$year<- as.numeric(emissions_US_Denmark$year)
+emissions_US_Denmark[1:30,2]<-1990:2019
+emissions_US_Denmark[31:60,2]<- 1990:2019
+```
+ str_replace
+ str_detect
+ str_split 
+ Feel free showing me the other way
+
+NA's were introduced by coercion if there was something other than a number entered in.
+
+## Checking Data for Outliers
+
+Lets check for outliers.
+```{r}
+summary(emissions_US_Denmark)
+```
+Even with both countries together the max is not far from the third quartile, so there appears to be no outliers. Lets check this with graphs.
+
+### Graphing
+Lets graph the two to see if there are any outliers.
+Lets use a scatterplot first, since we have emissions and year. 
+```{r}
+ggplot(emissions_US_Denmark, aes(x=year, y=emissions, colour=country)) + geom_point()+ theme_bw()
+```
+Eventually I would like to graph this with a line of best fit. How would you do that? Is it geom_smooth(method=lm), because it is not working.
+
+It looks like Denmark has significantly less emissions than the U.S.
+There are some odd spikes in Denmark's emissions, but nothing that seems to be an outlier.
+The data trends slowly downward which makes sense, given both countries are trying to reduce their emissions.
+
+Now time for a Boxplot
+
+Lets see with a boxplot how the data looks
+
+```{r}
+ggplot(emissions_US_Denmark, aes(x=country, y=emissions)) + geom_boxplot()+ theme_bw()
+```
+There standard deviation bars are not that big, espeically with the United States. The emissions look very reasonable for both.
+
+the next step will be to calculate rate and to do a test, but we will save that for after we clean up the other data sets.
+For tests we will do either a repeated measures analysis or a time series analysis.
+
+## Save the Corrected Data
+Saving corrected and reformatted emissions data.
+```{r}
+write.csv(emissions_US_Denmark, here("Data for the Code", "emissions_corrected_US_Denmark.csv"), row.names=FALSE)
+
+#row.names=FALSE prevents R from adding extra blank rows to the csv.
+
+write.csv(emissions_long, here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_all.csv"), row.names=FALSE)
+```
+
+## Working with the Facebook Data
+
+```{r}
+allfiles<- list.files(here("Data for the Code", "Facebook_data"), pattern="Facebook_2022")
+```
+
+```{r}
+pivot_longer(Agentina:Vietnam, names_to="Country", values_to="Number")
+```
 

### 2022-11-08 Code/Exploratory_Analysis_CAP_DAP.Rmd
facebook data working on it

@@
 
+Lets see one of the data sets. 
+```{r}
+DF<- read.csv(here("Data for the Code", "Facebook_data", "Facebook_2022_climate_awareness.csv"), stringsAsFactors = TRUE)
+```
+
+As of right now the Facebook data are in several different data csvs and all in wide format.
+
+We'd like to reformat these data so that every country is a row, the question is a new column, and the number responding with each answer is a new column.
+
+We'll use `pivot_longer` from `tidyr` package
+
+```{r}
+DF_long<- DF %>% pivot_longer(Albania:Zambia, names_to = "Country", values_to = "Number")
+
+write.csv(DF_long, here("Data for the Code", "Facebook_data", "ccos_awareness.csv"), row.names = FALSE)
+#Now we have the csv for the climate change opinion survey about awareness. 
+```
+
+That works well, however I would have to do it 30 times. So, lets do a loop that repeats this for each of the cc_awareness data sets?
+
+Get the list of files. Two of the files contain information just about the US and should not be used, so let's also exclude them.
+```{r}
+myfiles<-list.files(here("Data for the Code", "Facebook_data"),pattern ="Facebook_2022")
+
+```
+
+Now build the loop. Include code to strip the suffix of each file and add it to the output file name. It should also add a column with the question that the data are for.
+```{r}
+for (i in 1:length(myfiles)){
+  DF <- read.csv(here("Data for the Code", "Facebook_data", myfiles[i]))
+  DF_long<- DF %>% pivot_longer(Albania:Zambia, names_to = "Country", values_to = "Number")
+
+mystring<-str_split(myfiles[i], "Facebook_2022")
+
+Qname<-str_remove(mystring[[1]][2], pattern = ".csv")
+
+DF_long <- DF_long %>% mutate(question = paste0("Ques_", Qname))
+
+colnames(DF_long)[1] <- "response"
+
+my_filename<-paste0("ccos",mystring[[1]][2])
+
+write.csv(DF_long, here("Data for the Code", "Corrected_Data_sets", my_filename), row.names = FALSE)               
+}
+```
+Looks like it worked!
+
+Now let's open a few
+```{r}
+awareness<-read.csv(here("Data for the Code", "Corrected_Data_sets", "ccos_climate_awareness.csv"))
+
+beliefs <-read.csv(here("Data for the Code", "Corrected_Data_sets",  "ccos_Climate_beliefs.csv"))
+```
+Ok - that's all working. So now I can pull them all together into one big data frame.
+
+```{r}
+myfiles<-list.files(here("Data for the Code", "Corrected_Data_sets"),pattern ="ccos")
+
+newDF<-read.csv(here("Data for the Code", "Corrected_Data_sets", myfiles[1]))
+
+for (i in 2:length(myfiles)){
+  tempDF <- read.csv(here("Data for the Code", "Corrected_Data_sets", myfiles[i]))
+  newDF<-rbind(newDF, tempDF)
+}
+
+my_filename<-"ccos_all.csv"
+
+write.csv(newDF, here("Data for the Code", "Corrected_Data_sets",my_filename), row.names = F)
+
+```
+
+Now we can check for errors and outliers
+
+## Checking the Data
+
+suck it in
+```{r}
+ccos<-read.csv(here("Data for the Code", "Corrected_Data_sets", "ccos_all.csv"))
+```
+
+column names
+```{r}
+names(ccos)
+```
+We have the column names we were expecting spelled just fine. Lets further examine the data. 
+```{r}
+glimpse(ccos)
+```
+So we have here the resposnes, questions, and answers for every country in this survey.
+We can see the columns that are percent of people who answered that way are correctly labeled as double and not character data.
+We only want Denmark and the U.S. So, lets filter and see how the data looks.
+
+```{r}
+ccos_US_Denmark<- ccos %>% filter(Country=="United.States" | Country=="Denmark")
+```
+
+## Checking for outliers
+
+```{r}
+summary(ccos_US_Denmark)
+```
+This tells me it is the number of individuals who responded a certain way.
+The question is, why is the max 3096 and the third quartile is only 44.612?
+This big difference may indicate an outlier.
+The large number is the unweighted base of the survey population.
+What the data collectors did was "To ensure that the survey sample more accurately reflects the characteristics of the population, we assign weights to survey responses, which doesn’t identify anyone but helps correct for sample bias." So, the unweighted data is without that.
+We do not want to include that in our data.
+
+### Removing the unweighted base
+```{r}
+ccos_US_Denmark<- ccos_US_Denmark %>% filter(ccos_US_Denmark$ response!="(Unweighted Base)")
+```
+
+Lets run this again:
+
+```{r}
+summary(ccos_US_Denmark)
+```
+Now the 3rd quartile is 29.21 and the max is 88.19. this may be because the United States is larger than Denmark, skewing the means, so lets use graphs instead.
+
+```{r}
+ccos_US_Denmark_worry<- ccos_US_Denmark %>% filter(ccos_US_Denmark$question=="Ques__climate_worry")
+ggplot(ccos_US_Denmark, aes(x="response", y="Number")) + geom_bar(stat='identity', position='dodge', colour='Country')
+```
+
+To check for outliers I do not think we can graph all of them
+```{r}
+ggplot(ccos_US_Denmark, aes(x=Question))
+```
+
 ```{r}

### 2022-11-09 Code/Exploratory_Analysis_CAP_DAP.Rmd
Finishing up Facebook and Finishing final Dataset Yale

@@
 ```
-Now the 3rd quartile is 29.21 and the max is 88.19. this may be because the United States is larger than Denmark, skewing the means, so lets use graphs instead.
+Now the 3rd quartile is 29.21 and the max is 88.19.
+this may be because the United States is larger than Denmark, skewing the means, so lets use graphs instead.
+The high 88 appears to be in the question_happening section, so lets look at that.
 
 ```{r}
-ccos_US_Denmark_worry<- ccos_US_Denmark %>% filter(ccos_US_Denmark$question=="Ques__climate_worry")
-ggplot(ccos_US_Denmark, aes(x="response", y="Number")) + geom_bar(stat='identity', position='dodge', colour='Country')
+ccos_US_Denmark_happening<- ccos_US_Denmark %>% filter(ccos_US_Denmark$question=="Ques__happening") %>% group_by(Country)
+ggplot(ccos_US_Denmark_happening, aes(x=response, y=Number, fill=Country)) + geom_bar(stat='identity', position='dodge')+ scale_fill_manual(values=c(Denmark="light blue", United.States="light green"))+ theme(axis.text.x =element_text(angle = 90)) + theme_bw()
+#????The questions  on the x axis are blurred out, we should rotate them, which I attempted here and did not work, thoughts????
+```
+Both the U.S. and Denmark had a huge amount of people answer yes, so this is not an outlier, and thus our data is fine. Lets move on to our final data sheet, the Yale survey.
+
+```{r}
+write.csv(ccos_US_Denmark, here("Data for the Code", "Corrected_Data_sets", "final_ccos_US_Denmark.csv"), row.names = FALSE)
+```
+YAY
+
+## Yale Survey Data
+Bring in the data
+```{r}
+yale<- read.csv(here("Data for the Code", "Yale_International Perspective of Climate Change.csv"))
 ```
 
-To check for outliers I do not think we can graph all of them
+Lets start by checking out the column names.
+
 ```{r}
-ggplot(ccos_US_Denmark, aes(x=Question))
+names(yale)
+```
+Well this is most certainly not what we want.
+Though it makes sense, as the first row in the data merges the cells to be an overall header of seceral columns, but it is not the header of an individual column.
+So, we will have to delete that first row and hopefully that will fix the column names.
+
+Lets check the data further.
+
+```{r}
+glimpse(yale)
+```
+We have factor data for everything including ones that should be numerical data. We will need to change that back.
+
+What we need to fix:
+1. delete the first row
+2. fix column names
+3. make the columns doubles
+
+## Fixing the Yale Data
+
+### Deleting the First Row and trying to fix column names
+
+Wouldn't it be great if I could take the first row and make it a vector and then assign it as the column names? Well if I thought it, I am sure there is code for it.
+
+This was a pain
+```{r}
+as.vector(as.character((yale[1,])))
+#make a vector of the character data in row one of yale
+#??? not sure why this code is also needed up here, when I delted it, it did not work.???
+ccs_Yale_Corrected<- yale %>% slice(-1)
+#remove row one of yale and store it as a new data frame
+names(ccs_Yale_Corrected)<-c(as.vector(as.factor((yale[1,]))))
+#make the column names of the new data frame a vector of the character data in row one of yale
+```
+It did work though.
+
+Next issue.
+
+## Make the columns double.
+
+Unlike the data frame above, the columns that we need to switch to not all of them, but a select few. Lets see if : works. Pleaseeeee work.
+
+I tried it three different ways with "" and with $ and without both. None worked.
+
+```{r}
+indx <- sapply(ccs_Yale_Corrected, ccs_Yale_Corrected$Sample_2007:ccs_Yale_Corrected$Cl_aware)
+emissionsfix[indx] <- lapply(emissionsfix[indx], function(x) as.numeric(as.character(x)))
+```
+I will have to do a faster way later.
+
+By hand it works:
+```{r}
+ccs_Yale_Corrected$Aware<- as.numeric(ccs_Yale_Corrected$Aware)
 ```
 
+All right so Erika show me a faster way later. I am going to manually change the relevant ones to check my data out further.
+
+For this survey data I will be using the unaware and aware, and serious, and not serious data survey questions referring to climate change.
 ```{r}
-allfiles<- list.files(here("Data for the Code", "Facebook_data"), pattern="Facebook_2022")
+ccs_Yale_Corrected$Unaware<- as.numeric(ccs_Yale_Corrected$Unaware)
+ccs_Yale_Corrected$Serious<- as.numeric(ccs_Yale_Corrected$Serious)
+ccs_Yale_Corrected$Not_serious<- as.numeric(ccs_Yale_Corrected$Not_serious)
+
 ```
 
+## Checking for Outliers
 ```{r}
-pivot_longer(Agentina:Vietnam, names_to="Country", values_to="Number")
+summary(ccs_Yale_Corrected)
 ```
 
+The Not_serious data has more than 30 % difference from its third quartile to its maximum, so that needs to be further investigated.
+the unaware data should be double checked as well.
+The other two are free of outliers.
+
+## Graphing to Check Outliers
+
 ```{r}
+ggplot(ccs_Yale_Corrected, aes(x=Unaware)) + geom_histogram(binwidth = 10)+ theme_bw()
+```
 
+This is a fine spread with no apparent outliers.
+
+```{r}
+ggplot(ccs_Yale_Corrected, aes(x=Not_serious)) + geom_histogram(binwidth = 10)+ theme_bw()
 ```
+
+This histogram has a tail, but no outliers as well. The data looks good.
+
+!!!!We will save our finished Yale csv, so make sure to re-run this when all the columns are switched to numeric that should be.!!!!!
+
+## Saving the new dataframe
+```{r}
+write.csv(ccs_Yale_Corrected, here("Data for the Code","Corrected_Data_sets", "ccs_Yale_Corrected.csv"), row.names = FALSE)
+```

### 2022-11-14 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
fixing path changes exploratory

---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output: html_document
---

### 2022-11-14 Code/Final_Exploratory_Analysis_CAP_DAP.Rmd
Initial chunk of the final project for CAP DAP

--- 2022-11-09 Code/Exploratory_Analysis_CAP_DAP.Rmd
@@
 ```{r}
-emissions<-read.csv(here("Data for the Code", "CAIT_percapita_allcountries.csv"), stringsAsFactors = TRUE)
+emissions<-read.csv(here("Data for the Code", "Raw Data Sets", "CAIT_percapita_allcountries.csv"), stringsAsFactors = TRUE)
 ```
@@
 ```{r}
-DF<- read.csv(here("Data for the Code", "Facebook_data", "Facebook_2022_climate_awareness.csv"), stringsAsFactors = TRUE)
+DF<- read.csv(here("Data for the Code","Raw Data Sets", "Facebook_data", "Facebook_2022_climate_awareness.csv"), stringsAsFactors = TRUE)
 ```
@@
 
-write.csv(DF_long, here("Data for the Code", "Facebook_data", "ccos_awareness.csv"), row.names = FALSE)
+write.csv(DF_long, here("Data for the Code", "Raw Data Sets", "Facebook_data", "ccos_awareness.csv"), row.names = FALSE)
 #Now we have the csv for the climate change opinion survey about awareness. 
@@
 ```{r}
-myfiles<-list.files(here("Data for the Code", "Facebook_data"),pattern ="Facebook_2022")
+myfiles<-list.files(here("Data for the Code", "Raw Data Sets", "Facebook_data"),pattern ="Facebook_2022")
 
@@
 for (i in 1:length(myfiles)){
-  DF <- read.csv(here("Data for the Code", "Facebook_data", myfiles[i]))
+  DF <- read.csv(here("Data for the Code","Raw Data Sets", "Facebook_data", myfiles[i]))
   DF_long<- DF %>% pivot_longer(Albania:Zambia, names_to = "Country", values_to = "Number")
@@
 
-write.csv(DF_long, here("Data for the Code", "Corrected_Data_sets", my_filename), row.names = FALSE)               
+write.csv(DF_long, here("Data for the Code", "Corrected_Data_sets","Facebook csv's Long Format", my_filename), row.names = FALSE)               
 }
@@
 ```{r}
-awareness<-read.csv(here("Data for the Code", "Corrected_Data_sets", "ccos_climate_awareness.csv"))
+awareness<-read.csv(here("Data for the Code", "Corrected_Data_sets", "Facebook csv's Long Format", "ccos_climate_awareness.csv"))
 
-beliefs <-read.csv(here("Data for the Code", "Corrected_Data_sets",  "ccos_Climate_beliefs.csv"))
+beliefs <-read.csv(here("Data for the Code", "Corrected_Data_sets", "Facebook csv's Long Format",  "ccos_Climate_beliefs.csv"))
 ```
@@
 ```{r}
-myfiles<-list.files(here("Data for the Code", "Corrected_Data_sets"),pattern ="ccos")
+myfiles<-list.files(here("Data for the Code", "Corrected_Data_sets", "Facebook csv's Long Format"),pattern ="ccos")
 
-newDF<-read.csv(here("Data for the Code", "Corrected_Data_sets", myfiles[1]))
+newDF<-read.csv(here("Data for the Code", "Corrected_Data_sets","Facebook csv's Long Format", myfiles[1]))
 
 for (i in 2:length(myfiles)){
-  tempDF <- read.csv(here("Data for the Code", "Corrected_Data_sets", myfiles[i]))
+  tempDF <- read.csv(here("Data for the Code", "Corrected_Data_sets","Facebook csv's Long Format", myfiles[i]))
   newDF<-rbind(newDF, tempDF)
@@
 ```
-YAY
 
@@
 ```{r}
-yale<- read.csv(here("Data for the Code", "Yale_International Perspective of Climate Change.csv"))
+yale<- read.csv(here("Data for the Code", "Raw Data Sets", "Yale_International Perspective of Climate Change.csv"))
 ```
@@
 
-Wouldn't it be great if I could take the first row and make it a vector and then assign it as the column names? Well if I thought it, I am sure there is code for it.
+I could take the first row and make it a vector and then assign it as the column names.
 
-This was a pain
 ```{r}
@@
 #make a vector of the character data in row one of yale
-#??? not sure why this code is also needed up here, when I delted it, it did not work.???
+#??? not sure why this code is also needed up here, when I deleted it, it did not work.???
 ccs_Yale_Corrected<- yale %>% slice(-1)
@@
 ```
-It did work though.
 
@@
 
-Unlike the data frame above, the columns that we need to switch to not all of them, but a select few. Lets see if : works. Pleaseeeee work.
+Unlike the data frame above, the columns that we need to switch to not all of them, but a select few. Thus, sapply lapply will not work, so they must be changed individually.
 
-I tried it three different ways with "" and with $ and without both. None worked.
-
-```{r}
-indx <- sapply(ccs_Yale_Corrected, ccs_Yale_Corrected$Sample_2007:ccs_Yale_Corrected$Cl_aware)
-emissionsfix[indx] <- lapply(emissionsfix[indx], function(x) as.numeric(as.character(x)))
-```
-I will have to do a faster way later.
-
-By hand it works:
 ```{r}
@@
 ```
-
-All right so Erika show me a faster way later. I am going to manually change the relevant ones to check my data out further.
 
@@
 
-!!!!We will save our finished Yale csv, so make sure to re-run this when all the columns are switched to numeric that should be.!!!!!
-
 ## Saving the new dataframe

### 2022-11-14 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Initial chunk of the final project for CAP DAP

@@
 ---
+
+## Introduction
+
+In the great pursuit of mitigating climate change, countries have differed greatly in their strategies and successes.
+Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D.
+C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.] As other countries trail behind, illuminating how Denmark is successfully reducing its emissions is pertinent for other countries.
+Although reducing emissions involves extensive effort from multiple sectors, the citizens also greatly determine the success of mitigation plans.^[Weber, E.
+U., "What Shapes perceptions of climate change?," WIRES Climate Change.2010] What determines people's perceptions and willingness to decrease their own emissions, is therefore important for countries to understand.
+To delve into this question data from two global surveys on climate perceptions will be analyzed along with green house gas emissions and the reduction rate of countries.
+Special emphasis will be on comparing Denmark due to the countrie's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings.
+2022.]
+
+This data analysis will consist of three sections: the greenhouse gas emissions of the U.S.
+and Denmark, the climate change perceptions of their citizens, and the correlation between emissions and climate change perceptions, using data from many countries.
+
+In the emissions section Denmark's the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U.
+S.
+had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]
+
+The second section will include an analysis of several quesitons regarding climate change for both the U.S.
+and Denmark.
+These will whether the citizens are aware of climate change, how worried they are, and how big of a threat they believe climate change is in the next 20 years.
+We hypothesize Denmark will have a significantly higher percentage of people who are aware of climate change than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane.
+“Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502.
+DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more of Denmark's citizens will be very worried about climate change and see it as an extreme threat than U.
+S.
+citizens, because because climate change is uniformly depicted in Danish media as a problem, but it is not in the U.S.^4
+
+In the third section we will overlay the survey data with the emissions data to deterrmine the percentage of causation.
+We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang.
+C., " The influence of political ideology on greenhouse gas emissions" Global Environmental Change 74, no.1 (May 2022):102496.
+DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]
+
+## Statistcal Analysis and Results
+
+Statistical analysis will consist of the aforementioned three sections.
+
+## Climate Change Emissions of Countries
+
+Preparing R
+```{r}
+rm(list=ls())
+library(tidyverse)
+library(here)
+library(ggfortify)
+```
+
+Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.
+
+Lets first view the emissions of Denmark and the United States.
+
+Reading in the Data
+```{r}
+emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
+```
+
+### Creating the figure
+
+???HOW DO I GET RID OF THE LINE?
+```{r}
+
+#year <- function(x) as.POSIXlt(x)$year + 1900
+# I do not know how the above code does this :(
+ggplot(emissions, aes(year, emissions)) + 
+  geom_path() + geom_point(aes(colour=country))+ theme_bw()
+
+```
+### Running the Test
+```{r}
+
+```
+
+### Rate of Emissions
+
+To determine rate the emissions must be divided by the year.
+
+???? HOW DO THAT WITH TIME SERIES???
+
+## Perceptions of Climate Change
+
+## Summary Discussion
+
+## Challenges and Limitations

### 2022-11-14 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
added explanation sentence
survey data work

@@
 
-The second section will include an analysis of several quesitons regarding climate change for both the U.S.
+The second section will include an analysis of several questions regarding climate change for both the U.S.
 and Denmark.
-These will whether the citizens are aware of climate change, how worried they are, and how big of a threat they believe climate change is in the next 20 years.
-We hypothesize Denmark will have a significantly higher percentage of people who are aware of climate change than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane.
+These will include what Americans and Danes believe causeds climate change, how worried they are about it, and how big of a threat they believe climate change is in the next 20 years.
+We hypothesize Denmark will have a significantly higher percentage of people who are believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane.
 “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502.
-DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more of Denmark's citizens will be very worried about climate change and see it as an extreme threat than U.
+DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U.
 S.
-citizens, because because climate change is uniformly depicted in Danish media as a problem, but it is not in the U.S.^4
+citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not in the U.S.^4
 
-In the third section we will overlay the survey data with the emissions data to deterrmine the percentage of causation.
+In the third section we will overlay the survey data with the emissions data to determine the percentage of causation.
 We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang.
@@
 ```
+Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States.
+
 ### Running the Test
@@
 
+Reading in the data:
+```{r}
+survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
+```
+
+Though more people may have been intereviewed in country or the other, the survey data is in percent, as in the percentage of the whole who chose that option.
+We can double check to ensure the recorded numbers equals 100 per couontrry:
+```{r}
+total<- beliefs %>% group_by(Country) %>% summarise(total=sum(Number))
+```
+
+This means we can fairly compare Danes and American's perceptions of climate change.
+
+Lets first see what Danes and Americans believe caused climate.
+The options are the survey are: caused mostly by humans, caused by natural change in the environment, caused by both, and climate change is not happening.
+
+```{r}
+beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
+```
+
+```{r}
+ggplot(beliefs, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
+```
+More Americans believe that climate change is not happening or that it is caused by natural changes than Danes.
+More Danes believe it is caused by human activities.
+However, more Danes than Americans think it is cause equally by both.
+Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.
+
+We have multiple categorical predictor variables with ratio data as the response variable, so we will use an ANOVA.
+
+generalized linear model. Chapter 7 in my book.
+
+```{r}
+
+```
+
 ## Summary Discussion

### 2022-11-16 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
working on general linear model

@@
 
-Though more people may have been intereviewed in country or the other, the survey data is in percent, as in the percentage of the whole who chose that option.
-We can double check to ensure the recorded numbers equals 100 per couontrry:
+Though more people may have been interviewed in one country than the other, the survey data is in percent, as in the percentage of the whole who chose that option.
+We can double check to ensure the recorded numbers equals 100 per country per survey question.
+
+Lets check with the 'beilefs question' which is the first we will anaylzed.
+
+```{r}
+#Parcing out just the survey answers regarding climate change beliefs
+beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
+```
+
 ```{r}
@@
 
+### Beliefs of Climate Change
+
 Lets first see what Danes and Americans believe caused climate.
@@
 ```{r}
-beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
-```
-
-```{r}
-ggplot(beliefs, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
+# Making the graph
+ggplot(beliefs, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
+  xlab("Response to Survey Question")+ ylab("Percent") + # renaming x and y axis labels
+  theme_bw() +
+  coord_flip() # switching x and y axis to read the suvey questions
 ```
@@
 
-We have multiple categorical predictor variables with ratio data as the response variable, so we will use an ANOVA.
+```{r}
+ggplot(beliefs, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+The pie chart clearly shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.
 
-generalized linear model. Chapter 7 in my book.
+We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a generalized linear model.
+
+### Running the Statistical Test
+
+The family is not binoomial, poisson, or gamma, ugh what is it?
 
 ```{r}
+glm(Number~response, data=beliefs, family=) #Constructig the general linear model
+#The family (or ‘error’) part of a GLM is not too difficult to understand.
+This bit just determines what kind of distribution is used to describe the response variable.
+A general linear model always assumes normality— that’s pretty much its defining feature in fact—but with a generalized linear
+```
 
+### Level of worriedness about Climate Changed
+In this survey question people were asked how worried they are about climate change. There are three options for 
+```{r}
+worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
 ```
+
+```{r}
+ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
+```
+32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.
+
+```{r}
+ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+
+Lets do a statistical test to verify.
 

### 2022-11-17 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Did as much as I could on survey data

@@
 library(ggfortify)
+library(forecast)
 ```
@@
 ```{r}
-emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
+emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"))
 ```
+
+Lets turn the date column into date data using the lubridate package.  
+To do that we need to put in the day and the month using string.
+
+```{r}
+emissions$year<-paste0("01-01-", emissions$year)
+```
+
+???This NA'd all my data and %y messed my dates up so I had to do the code below???
+```{r}
+emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
+#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
+# must do the - in the %d- because the dates are entered with a -. 
+```
+
+#Now when running statistical tests R will know the data are dates.
 
@@
 ### Running the Test
+
+First determine the minimum date
 ```{r}
+min_date=min(emissions$year)
+max_date=max(emissions$year)
+```
 
+Turning our data into a time series
+
+```{r}
+emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
 ```
+
+```{r}
+head(emissions_ts)
+plot(emissions_ts, xlab="year", ylab="emissions")
+```
+Making a moving average:
+```{r}
+Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
+```
+
+```{r}
+plot(emissions_ts)
+lines(Ma_emissions, col="maroon", lwd=3)
+```
+Not sure if I am supposed to do this next part...
+
+```{r}
+seasonal_add<-emissions_ts - Ma_emissions
+plot(seasonal_add)
+```
+ya.. this looks weird.
+
+```{r}
+fit_add<-decompose(emissions_ts, type='additive')
+plot(fit_add)
+```
+Ya, so I cannot decompose my data. She also did not teach us how to run the test, which is annoying.
 
@@
 
-The family is not binoomial, poisson, or gamma, ugh what is it?
+The family is not binomial, poisson, or gamma, ugh what is it? It accepted gaussain but autoplot looked wrong.
+
+Gaussian family : for continuous decimal data with normal distribution, like weight, length, et al.
+frequency.
+So given I have percentage data, which is similar, I believe I should use poisson.
+However,
 
 ```{r}
-glm(Number~response, data=beliefs, family=) #Constructig the general linear model
+belief_glm<-glm(Number~response, data=beliefs, family= Gamma(link='inverse') ) #Constructing the general linear model
 #The family (or ‘error’) part of a GLM is not too difficult to understand.
@@
 A general linear model always assumes normality— that’s pretty much its defining feature in fact—but with a generalized linear
+
+#changing the default link yoou doo family=poiossoono(link=log)) for log.
+Remember, this is the log link function for Poisson models.
+If you want to be explicit about the link function, or change the default, this
+
 ```
+Test autoplot
+
+```{r}
+autoplot(belief_glm)
+```
+Something is still wrong
+
+```{r}
+anova(belief_glm)
+```
+???I believe I am not choosing the right family=.??? 
+???? Or I have really underdispersed data?
 
@@
 ```
+We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all.
+Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%).
+More Americans however are very worried, and more Americans are not at all worried than Danes.
+Interesting.
+A statistical test is needed to see if Americans are less worried than Danes.
 
-Lets do a statistical test to verify.
+### Lets do a statistical test to verify. 
+Again we will do a general linear model.
+
+```{r}
+worry_glm<-glm(Number~response, data=worry, family= Gamma(link='inverse') ) #Constructing the
+
+```
+Test autoplot
+
+```{r}
+autoplot(worry_glm)
+```
+again, very concerning
+
+```{r}
+anova(worry_glm)
+```
+## Is Climate Change a Big Threat
+
+In this survey question people were asked whether or not climate change is a big threat within the next 20 years.
+The options were very serious threat, somewhat serious, not a threat at all, or I do not know.
+
+parcing the data
+```{r}
+threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
+```
+
+```{r}
+ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
+```
+More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans.
+More Americans also think that cliamte change is not a threat at all than Danes.
+Lets view a pie chart.
+
+```{r}
+ggplot(threat, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat.
+Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that.
+Lets test with a statistical test.
+
+### Lets do a statistical test to verify. 
+Again we will do a general linear model.
+
+```{r}
+threat_glm<-glm(Number~response, data=threat, family= Gamma(link='inverse') ) #Constructing the
+
+```
+Test autoplot
+
+```{r}
+autoplot(threat_glm)
+```
+again, very concerning
+
+```{r}
+anova(threat_glm)
+```
+## More or less Fossil Fuels
+
+Now that we know what Danes and Americans think about climate change and how worried they are, lets see how they are reacting.
+This next survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.
+
+Lets check with the 'beilefs question' which is the first we will anaylzed.
+
+```{r}
+#Parcing out just the survey answers regarding climate change beliefs
+fuels<- survey %>% filter(survey$question=="Ques__Fossil_moreorless") 
+```
+
+```{r}
+# Making the graph
+ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
+  xlab("Response to Survey Question")+ ylab("Percent") + # renaming x and y axis labels
+  theme_bw() +
+  coord_flip() # switching x and y axis to read the suvey questions
+```
+More Americans think there should be much more or or somewhat more fossil fuels used than Danes.
+More Danes thinks threre should be somewhat less or much less fossil fuels used.
+This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less.
+Lets see in pie char form.
+
+```{r}
+ggplot(fuels, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+The pie chart clearly shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used.
+These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%.
+More Americans believed there should be much more or somewhat more than Danes.
+Lets verify with a statistical test.
+
+we will run another generalized linear model.
+
+### Running the Statistical Test
+
+```{r}
+fuels_glm<-glm(Number~response, data=fuels, family= Gamma(link='inverse') ) #Constructing the 
+```
+Test autoplot
+
+```{r}
+autoplot(fuels_glm)
+```
+Something is still wrong
+
+```{r}
+anova(fuels_glm)
+```
+???I believe I am not choosing the right family=.??? 
+???? Or I have really underdispersed data?
+
+## Survey data Impacting Emissions
 

### 2022-11-17 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Changes w/ Erika

@@
 # I do not know how the above code does this :(
-ggplot(emissions, aes(year, emissions)) + 
-  geom_path() + geom_point(aes(colour=country))+ theme_bw()
+ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  geom_line() +
+  geom_point(aes(colour=country))+ theme_bw()
 
@@
 ```{r}
-belief_glm<-glm(Number~response, data=beliefs, family= Gamma(link='inverse') ) #Constructing the general linear model
+beliefs<- beliefs %>% mutate(frequency=Number/100)
+belief_glm<-glm(frequency~response, data=beliefs, family= binomial) #Constructing the general linear model
 #The family (or ‘error’) part of a GLM is not too difficult to understand.
@@
 ```
+
+```{r}
+ggplot(beliefs, aes(frequency, group=Country, fill=Country, position='dodge'))+ geom_histogram()+theme_bw()
+```
+
 Test autoplot
@@
 ```{r}
-ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
+ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
 ```

### 2022-11-18 Code/Draft_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
created cleaned up final version

--- 2022-11-17 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
@@
 ```{r}
-fit_add<-decompose(emissions_ts, type='additive')
-plot(fit_add)
+#fit_add<-decompose(emissions_ts, type='additive')
+#plot(fit_add)
 ```
@@
 More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans.
-More Americans also think that cliamte change is not a threat at all than Danes.
+More Americans also think that climate change is not a threat at all than Danes.
 Lets view a pie chart.
@@
 
+### Viewing the Survey Data
+lets view all the survey data in the same graph with a likert graph. 
+I need my data in wide format based on response, so I need all four to have the same response.
+None of them have the same response, and even if I changed the responses they are not in the right order to do.
+It would be alot of work to do this, so I am going to keep the pie chart method.
+```{r}
+questions<- survey %>% filter(survey$question=="Ques__Climate_beliefs" | survey$question=="Ques__climate_worry" | survey$question=="Ques__threat_20_years" | survey$question=="Ques__Fossil_moreorless" )
+
+new_q<- c("very much","some", "nuetral", "other", "harms","refuse")
+questions_new<- questions %>% filter(survey$question=="Ques__Climate_beliefs") %>% mutate(response=
+
+#survey_wide <– quesitons, id_cols=
+```
+
+```{r}
+library(likert)
+plot(likert(questions), ordered = F, wrap= 60)
+```
+
 ## Survey data Impacting Emissions

### 2022-11-27 Code/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
feedback on data exploration and draft analysis

---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output: html_document
---

## Introduction

In the great pursuit of mitigating climate change, countries have differed greatly in their strategies and successes. Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D. C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.] As other countries trail behind, illuminating how Denmark is successfully reducing its emissions is pertinent for other countries. Although reducing emissions involves extensive effort from multiple sectors, the citizens also greatly determine the success of mitigation plans.^[Weber, E. U., "What Shapes perceptions of climate change?," WIRES Climate Change.2010] What determines people's perceptions and willingness to decrease their own emissions, is therefore important for countries to understand. To delve into this question data from two global surveys on climate perceptions will be analyzed along with green house gas emissions and the reduction rate of countries. Special emphasis will be on comparing Denmark due to the countrie's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings. 2022.]

I might use active voice "I will analyze data" vs. "data will be analyzed"

This data analysis will consist of three sections: the greenhouse gas emissions of the U.S. and Denmark, the climate change perceptions of their citizens, and the correlation between emissions and climate change perceptions, using data from many countries.

In the emissions section Denmark's and the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]

Go ahead and use "I" rather than "We" because this is  your work.

The second section will include an analysis of several questions regarding climate change for both the U.S. and Denmark. These will include what Americans and Danes believe caused climate change, how worried they are about it, how big of a threat they believe climate change is in the next 20 years, and if they believe fossil fuel usage should decrease, increase, or stay the same. We hypothesize Denmark will have a significantly higher percentage of people who believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane. “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502. DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U. S. citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4 Lastly, we hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)

In the third section we will overlay the survey data with the emissions data to determine the percentage of causation. We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang. C., " The influence of political ideology on greenhouse gas emissions" Global Environmental Change 74, no.1 (May 2022):102496. DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]

Good overall, but also include links to each dataset.

## Preparing R
```{r, message=FALSE}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(forecast)
library(gridExtra)
```

## Statistcal Analysis and Results

Statistical analysis will consist of the aforementioned three sections.

### Climate Change Emissions of Countries

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.

Lets first view the emissions of Denmark and the United States.

Reading in the Data
```{r}
emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"))
```

Lets turn the date column into date data using the lubridate package.  
To do that we need to put in the day and the month using paste.

```{r}
emissions$year<-paste0("01-01-", emissions$year)
```

Now tell R the year column are dates so we can better perform statistical analysis.

Below works fine, but just so you are aware, it is base R and not lubridate.

```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

Now when running statistical tests R will know the data are dates.

#### Creating the figure

Lets see the emissions of Denmark and the United States from 1990 to 2019.
```{r}
ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.

```

Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States. Both countries have been decreasing their emissions since 1990, but  Denmark seems to have a  greater rate of decrease in emissions, especially since 2006, than the United States.

Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.

I talked to Robin Lock about this. We're going to do a paired t-test where each year has a pair of values (US, Denmark) and the values will be the RATE of change for that year, rather than the GHG emissions themselves.

Here's some info: https://www.statology.org/paired-samples-t-test-r/

We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series). This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.

#### Running the Test

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```
I'm not sure you'll need the above code chunk

Turning our data into a time series

```{r}
emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Lets do time series for the indiviual countries
```{r}
emissions_Denmark<- emissions %>% filter(country=="Denmark")
emissions_US<-emissions %>% filter(country=="United States")
emissions_Denmark_ts<-ts(emissions_Denmark$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
emissions_US_ts<-ts(emissions_US$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Viewing the time series of each individual country to ensure its correct.
```{r, results='hide'}
both_ts<-cbind(emissions_Denmark_ts, emissions_US_ts)
head(both_ts) # ensuring it worked
plot(both_ts, xlab="year", ylab="emissions")
```

If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:

```{r}
demo_plot<-autoplot(emissions_US_ts)
demo_plot+
  labs(x = "Year", y = "Greenhouse gas emissions")+
  theme_bw()
```

And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:

```{r}
both<-cbind(emissions_Denmark_ts, emissions_US_ts)
autoplot(both)
```

Here is our data smoothed over so we can see the trends without added noise. The bottom one I presume is denmark. We can see that it is decreasing more.

Making a moving average:
```{r}
Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
Ma_emissions_Denmark<- forecast:: ma(emissions_Denmark_ts, order=3, centre=TRUE)
Ma_emissions_US<- forecast::ma(emissions_US_ts, order=3, centre=TRUE)

```

Plotting both the moving average and the time series.

```{r, collapse=TRUE}
plot(emissions_ts)
lines(Ma_emissions, col="maroon", lwd=3)

plot(emissions_Denmark_ts)
lines(Ma_emissions_Denmark, col="orange", lwd=3)

plot(emissions_US_ts)
lines(Ma_emissions_US, col="blue")
```
Plotting both with autoplot/ggplot:

```{r}
USA<-cbind(emissions_US_ts, Ma_emissions_US)
autoplot(USA)
```

The red line is the data but smoothed, so we can see the overall trend of the emissions data combined. the other two are them separated, notice how similar Denmark's is to the combined data. Interesting.

***PUT RUNNING THE TIME SERIES HERE***
This will really be the paired t-test

### Perceptions of Climate Change

Reading in the data:
```{r}
survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
```

Though more people may have been interviewed in one country than the other, the survey data is in percent, as in the percentage out of the whole who chose that option. We can double check to ensure the recorded numbers equals 100 per country per survey question.

Lets check with the 'beliefs question' which is the first question we will analyzed.

```{r}
#Parsing out just the survey answers regarding climate change beliefs
beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
```

```{r}
total<- beliefs %>% group_by(Country) %>% summarise(total=sum(Number))
total
```

This means we can fairly compare Danes and American's perceptions of climate change.

I talked with Prof Shuckers about these data. He thinks we should do the back-calculating to change percents to numbers and then do a chi-square to compare answers between Danes and Americans.  You will need the data on number of respondants for each question to do that.

#### Beliefs of Climate Change

Lets first see what Danes and Americans believe caused climate. The options to choose from were: caused mostly by humans, caused by natural change in the environment, caused by both, and climate change is not happening.

```{r}
# Making the graph
ggplot(beliefs, aes(response,Number, fill=Country))+
  geom_bar(stat = 'identity', position = 'dodge') + 
 labs(title="Cause of Climate Change", x="Response to Survey Question", y="Percent") + # titling and renaming x and y axis labels
  #theme( axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))+ this won't work :(
  coord_flip()+
  theme_bw()

```

Figure 1. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.

More Americans believe that climate change is not happening or that it is caused by natural changes than Danes. More Danes believe it is caused by human activities. However, more Danes than Americans think it is cause equally by both. Lets look at a pie chart as well.

If you want to test these questions with a statistical test, we will need a combination of chi-square tests. Let's talk.

```{r}
piecolor<- c("orange","maroon","darkgreen","lightblue", "gray70","purple")

beliefs<- beliefs %>% arrange(desc(Number)) %>% mutate(Position=cumsum(Number/2)-0.5*Number)
legend<- c("caused equally","caused mostly by humans","caused mostly by natural changes","climate change isin't happening","other","refused") #see below
ggplot(beliefs, aes(x="", Percent, y= Number, fill=response))+  
  geom_bar(stat="identity", width=1, color="white")+
  coord_polar("y", start=0) + 
  facet_wrap(~Country)+
 #legend( 3, legend=legend)+, attempt to make the legends smaller, so the pie charts could be bigger and I could move the percentages.
  scale_fill_manual(values=piecolor )+
   theme_void()+# remove background, grid, numeric labels
  geom_label(aes (label = round(Number, digits = 1), x=1.7),color="white", size=4)  # add percentages

```

Figure 2. Pie chart of what Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.

Here's an article that talks more about data viz for likert type data. It has good arguments for why to avoid pie charts, and has a building series of charts ending with the author's (and my) favorite. Just worth considering:

https://nightingaledvs.com/visualizing-likert-scale-data-same-data-displayed-seven-different-ways/

The pie chart shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.

Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.

We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a _________

Chi-square test after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.

***INSERT ### Running the Statistical Test***

#### Level of worriedness about Climate Changed
In this survey question people were asked how worried they are about climate change. There are three options for 
```{r}
worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
```

```{r}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity') + 
  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

Figure 3. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.

```{r}
ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 4. Pie chart of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all. Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%). More Americans however are very worried, and more Americans are not at all worried than Danes. Interesting. A statistical test is needed to see if Americans are less worried than Danes.

#### Lets do a statistical test to verify. 
Chi-square test. And think about a new data vis.

***### INSERT STATISTICAL TEST***

### Is Climate Change a Big Threat

In this survey question people were asked whether or not climate change is a big threat within the next 20 years. The options were very serious threat, somewhat serious, not a threat at all, or I do not know.

parsing the data
```{r}
threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
```

```{r}
ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Perceieved Threat of Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

Figure 5. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans. More Americans also think that climate change is not a threat at all than Danes. Lets view a pie chart.

```{r}
ggplot(threat, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.5, nudge_y = 2 ) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 6. Pie char of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat. Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that. Lets test with a statistical test.

#### Lets do a statistical test to verify.

***###INSERT STATISTICAL TEST ***

### More or less Fossil Fuels

Now that we know what Danes and Americans think about climate change and how worried they are, lets see how they are reacting. This next survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.

Lets check with the 'beilefs question' which is the first we will anaylzed.

```{r}
#Parcing out just the survey answers regarding climate change beliefs
fuels<- survey %>% filter(survey$question=="Ques__Fossil_moreorless") 
```

```{r}
# Making the graph
ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```

Figure 7. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.

More Americans think there should be much more or or somewhat more fossil fuels used than Danes. More Danes thinks threre should be somewhat less or much less fossil fuels used. This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less. Lets see in pie char form.

```{r}
ggplot(fuels, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 8. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.

The pie chart shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used. These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%. More Americans believed there should be much more or somewhat more than Danes. Lets verify with a statistical test.

we will run another generalized linear model.

#### Running the Statistical Test

***### INSERT STATISTICAL TEST***

## Survey data Impacting Emissions

## Summary Discussion

## Limitations
The data used is from a survey conducted on Facebook, so all respondents are Facebook users. This adds a confounding variable, when utilizing this study.

## Challenges

Prior to this project how to conduct a time series analysis, work with survey data, or make pie charts was not discussed nor taught. External research and problem solving had to be conducted in order to achieve this data analysis.

What about challenges in terms of figuring out how to reshape/reformat your data? Anything worth mentioning there?

This is very thorough and a solid first draft.

When constructing graphs the pie charts proved to be particularly cumbersome as the percentages were not align on the pie chart in the correct spot, nor all be visible. Research into geom_label and geom_text yeilded a the most succesful outcome. However, the pie charts still need work and any feedback regarding how to would be greatly appreciated.

Research into how to make a likert graph was also conducted. The responses to the four survey questions would have to be the same however, in order to produce a likert graph. I could change the answers to create a made up response scale of pro or anti climate action. Would that be professional or no? Let me know your thoughts so I can include it in the final.

### 2022-11-27 Code/Final_Exploratory_Analysis_CAP_DAP.Rmd
feedback on data exploration and draft analysis

@@
 
-## Emitting the Last Two Rows
+## Omitting the Last Two Rows
 ```{r}
@@
 Lastly, we need to make all the date columns into numeric rather than factor data, so when we put it in long format R knows they are numbers.
+
+If you dropped `stringsAsFactor = T` when you read in the data you can save yourself a big headache here. There are two issues:
+
+1. R won't allow column names that are numbers so it adds the X. This does not change the data type to factor- but it makes it difficult to use the numeric information in the dates.
+
+2.
+Because the folks who created the data used FALSE in some of the year columns, instead of NA (which R recognizes), R interprets those columns as character data (or, in this case, factor, because you imported with stringsAsFactors = T)
 
@@
 ```
+Well done! I'm not sure this was the easiest route, but you figured it out!
+
 NA's were introduced by coercion into the countries that have FALSE listed as their emissions, because their emissions could not be recorded.
@@
 ```
+The trouble with your way is that it cannot be automated, and is thus more prone to user error. It would also be increasingly more difficult with larger and larger data sets.
+
+You could try this (my demo is with emissions_long)
+
+```{r}
+erika_demo<- emissions_long
+erika_demo$year<-str_replace(erika_demo$year, "X", "")
+erika_demo$year<-as.numeric(erika_demo$year)
+```
+
  str_replace
@@
  Feel free showing me the other way
+ See above
 
@@
 
+It works when I do it - not sure the issue. And I'm not sure you want method = "lm" because these are not linear plots.
+
 It looks like Denmark has significantly less emissions than the U.S.
@@
 
+I wonder if it would be interesting to find the population size of USA and Denmark in all of those years and then calculate emissions per capita - Denmark will always have lower emissions than the US because it is a smaller country.
+
+Another option, which I discussed with professor Lock in Stats, is to look at the RATE of emissions change from year to year since the rate of change is independent of population size.
+
+Think about how you would calculate the rate of change.
 Now time for a Boxplot
@@
 
+How many of the Facebook questions are you going to examine?
+
 ## Yale Survey Data
@@
 
+You have character data you mean, not factor.
+
 What we need to fix:
@@
 #make the column names of the new data frame a vector of the character data in row one of yale
+```
+You made this more complicated than it needed to be, but it worked. Here's another, easier way.
+
+```{r}
+erika_demo2 <- yale
+goodNames<-slice(erika_demo2, 1) %>% unlist(use.names = FALSE)
+names(erika_demo2)<-goodNames
+erika_demo2<-erika_demo2[2:nrow(erika_demo2),]
 ```

### 2022-11-28 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
moving files making draft folder

--- 2022-11-27 Code/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
@@
 
-I might use active voice "I will analyze data" vs. "data will be analyzed"
-
 This data analysis will consist of three sections: the greenhouse gas emissions of the U.S.
@@
 
-In the emissions section Denmark's and the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U.
+In the emissions section Denmark's the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U.
 S.
 had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]
-
-Go ahead and use "I" rather than "We" because this is  your work.
 
@@
 DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]
-
-Good overall, but also include links to each dataset.
 
@@
 Now tell R the year column are dates so we can better perform statistical analysis. 
-
-Below works fine, but just so you are aware, it is base R and not lubridate.
-
 ```{r}
@@
 Both countries have been decreasing their emissions since 1990, but  Denmark seems to have a  greater rate of decrease in emissions, especially since 2006, than the United States.
-
 Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.
-
-I talked to Robin Lock about this.
-We're going to do a paired t-test where each year has a pair of values (US, Denmark) and the values will be the RATE of change for that year, rather than the GHG emissions themselves.
-
-Here's some info: https://www.statology.org/paired-samples-t-test-r/
-
-We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series).
-This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.
 
@@
 ```
-I'm not sure you'll need the above code chunk
 
@@
 plot(both_ts, xlab="year", ylab="emissions")
-```
-
-If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:
-
-```{r}
-demo_plot<-autoplot(emissions_US_ts)
-demo_plot+
-  labs(x = "Year", y = "Greenhouse gas emissions")+
-  theme_bw()
-```
-
-And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:
-
-```{r}
-both<-cbind(emissions_Denmark_ts, emissions_US_ts)
-autoplot(both)
 ```
@@
 ```
-Plotting both with autoplot/ggplot:
-
-```{r}
-USA<-cbind(emissions_US_ts, Ma_emissions_US)
-autoplot(USA)
-```
 
@@
 ***PUT RUNNING THE TIME SERIES HERE***
-This will really be the paired t-test
 
@@
 
-I talked with Prof Shuckers about these data.
-He thinks we should do the back-calculating to change percents to numbers and then do a chi-square to compare answers between Danes and Americans.
-You will need the data on number of respondants for each question to do that.
-
 #### Beliefs of Climate Change
@@
 
-If you want to test these questions with a statistical test, we will need a combination of chi-square tests. Let's talk.
-
 ```{r}
@@
 
-Here's an article that talks more about data viz for likert type data.
-It has good arguments for why to avoid pie charts, and has a building series of charts ending with the author's (and my) favorite.
-Just worth considering:
-
-https://nightingaledvs.com/visualizing-likert-scale-data-same-data-displayed-seven-different-ways/
-
 The pie chart shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.
@@
 We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a _________
-
-Chi-square test after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.
 
@@
 #### Lets do a statistical test to verify. 
-Chi-square test. And think about a new data vis.
+Again we will do a general linear model.
 
@@
 
-parsing the data
+parcing the data
 ```{r}
@@
 
-What about challenges in terms of figuring out how to reshape/reformat your data? Anything worth mentioning there?
-
-This is very thorough and a solid first draft.
-
 When constructing graphs the pie charts proved to be particularly cumbersome as the percentages were not align on the pie chart in the correct spot, nor all be visible.

### 2022-11-28 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
superscripts

@@
 date: "11/13/2022"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 S.
-citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4 Lastly, we hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)
+citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, we hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^
 

### 2022-11-29 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Enough to make a professor swear

@@
 ```{r}
-emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"))
+emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
 ```
 
-Lets turn the date column into date data using the lubridate package.  
+Lets turn the date column into date data using base R.  
 To do that we need to put in the day and the month using paste.
@@
 Now tell R the year column are dates so we can better perform statistical analysis.
-
-Below works fine, but just so you are aware, it is base R and not lubridate.
 
@@
 Here's some info: https://www.statology.org/paired-samples-t-test-r/
+
+```{r}
+emissions$rate<-NA
+denmark_rate<- emissions %>% filter(country=="Denmark")
+den_rate<- diff(denmark_rate$emissions)
+denmark_rate$rate[2:30]<-den_rate
+denmark_rate<-denmark_rate[2:30,]
+
+emissions$rate<-NA
+usa_rate<- emissions %>% filter(country=="United States")
+us_rate<- diff(usa_rate$emissions)
+usa_rate$rate[2:30]<-us_rate
+usa_rate<-usa_rate[2:30,]
+
+emissions_rate<- rbind(usa_rate, denmark_rate)
+emissions_rate_2010<- emissions_rate %>% filter(emissions_rate$year>="2005-01-01")
+
+```
+
+```{r}
+ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
+  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
+```
+
+```{r}
+t.test(rate~country, data=emissions_rate, paired=TRUE)
+#complete.cases means just use the ones without NAs
+```
+
+```{r}
+t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
+```
 
@@
 You will need the data on number of respondants for each question to do that.
+
+???______
+1101 people answered in Denmark
+3096 answered in United States
+
+```{r}
+beliefs <- beliefs %>% group_by(Country)
+
+dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+beliefs<- cbind(beliefs, dumb_way)
+beliefs<-beliefs %>% rename(num_people=...5)
+
+beliefs<- beliefs %>% group_by(Country, response) %>%  mutate(total=(Number/100)*num_people)
+
+beliefs$total<- round(beliefs$total, digits=0)
+
+test<- beliefs %>% select(response, Country, total)
+```
+
+???? The chi-squared below did not work!!!
+Chi-squared
+```{r}
+beliefs_matrix<-xtabs(total~Country + response, data=test)
+beliefs_matrix
+beliefs_df<- as.data.frame(beliefs_matrix)
+beliefs_df<- beliefs_df %>% filter(Freq>0)
+beliefs_matrix_real<- xtabs(Freq~Country+response, data=beliefs_df)
+beliefs_matrix_real
+chisq.test(beliefs_matrix)
+
+```
+???___
+
+```{r}
+vector_Den<- c(585,57,419,19,16,1)
+vector_US<-c(1295,473,1073,86,161,7)
+matrix<- as.matrix(rbind(vector_Den, vector_US))
+chisq.test(matrix)
+
+```
 

### 2022-11-30 Code/Drafts/Final_Short_Oral_Presentation.Rmd
Oral Presentation Preparation

---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output:
  pdf_document: default
  html_document: default
---

## Introduction

> ***Comparative study of U.S. and Denmark***

>Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D. C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.]

### What Data To Use?

>CAIT Data by (World Resource Institute) WRI from 1990-2019 including all emissions and per capita

>Survey Data 2022 from Yale on Facebook, was in wide format but had more questions

### Questions Asked:

### Emissions

#### 1. Does Denmark have less emissions per capita than the U.S.
#### 2. which has a faster rate of decrease in per capita GHG emissions

### Survey Data

#### 1. believe caused climate change

#### 2. how worried they are

#### 3. how big of a threat they believe it is in the next 20 years, and

### 4. If they believe fossil fuel usage should decrease, increase, or stay the same.

*We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]*

*I hypothesize Denmark will have a significantly higher percentage of people who believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane. “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502. DOI: 10.22230/cjc.2011v36n3a2367] I further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U. S. citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, we hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^*

## Preparing R
```{r, message=FALSE}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(forecast)
library(gridExtra)
```

## Statistcal Analysis and Results

Statistical analysis will consist of the aforementioned three sections.

### Climate Change Emissions of Countries

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.

Lets first view the emissions of Denmark and the United States.

Reading in the Data
```{r}
emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", 
  "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
```

Lets turn the date column into date data using base R.  
To do that we need to put in the day and the month using paste.

```{r, echo=FALSE}
emissions$year<-paste0("01-01-", emissions$year)
```

Now tell R the year column are dates so we can better perform statistical analysis.

```{r, echo=FALSE}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, 
#origin informs R where it starts. must do the - in the %d- because the dates 
#are entered with a -. 
```

Now when running statistical tests R will know the data are dates.

#### Creating the figure

Lets see the emissions of Denmark and the United States from 1990 to 2019.
```{r}
ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  
  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() 
#colour colors by country, and group tells R to group the points by country
#so geom_line will draw lines connecting the data by country.

```
Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States. Both countries have been decreasing their emissions since 1990, but  Denmark seems to have a  greater rate of decrease in emissions, especially since 2006, than the United States. A test almost seems silly, but lets run one to make sure.
```{r}
t.test(emissions~country, data=emissions, paired=TRUE)
```
The p-value is very small, much less than 0.05, so we can reject our Null hypothesis. Denmark does have signifacantly less GHG emissions per capita

#### What about the Rate of Decrease?

Calculating rate
```{r}
emissions$rate<-NA
denmark_rate<- emissions %>% filter(country=="Denmark")
den_rate<- diff(denmark_rate$emissions) # subtracts n by n-1 to get rate.
denmark_rate$rate[2:30]<-den_rate
denmark_rate<-denmark_rate[2:30,]

emissions$rate<-NA
usa_rate<- emissions %>% filter(country=="United States")
us_rate<- diff(usa_rate$emissions)
usa_rate$rate[2:30]<-us_rate
usa_rate<-usa_rate[2:30,]

emissions_rate<- rbind(usa_rate, denmark_rate)

```

Lets look at this smoothed over.

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
#group the points by country so geom_line will draw lines connecting the 
#data by country.

```

#### Running the Test
Lets run a paired t test

```{r, results='hold', echo=FALSE}
t.test(rate~country, data=emissions_rate, paired=TRUE) 
#complete.cases means just use the ones without NAs
```
A p value of 0.7668, so we have to accept our null hypothesis.

Lets determine if there is a difference in rate from when both country's emissions started significantly decreasing and the rate is consistently negative 2010.
```{r, echo=FALSE}
emissions_rate_2010<- emissions_rate %>% 
  filter(emissions_rate$year>="2010-01-01")

```

```{r, results='hold', echo=FALSE}
t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
#complete.cases means just use the ones without NAs
```
P value of 0.083, so we must again accept our null hypothesis.

### Perceptions of Climate Change

Reading in the data:
```{r}
survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", 
                       "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
```

### More or less Fossil Fuels

This survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.

```{r,echo=FALSE}
#Parcing out just the survey answers regarding climate change beliefs
fuels<- survey %>% filter(survey$question=="Ques__Fossil_moreorless") 
```

```{r, warning=FALSE, echo=FALSE}
# Making the graph
ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
  position = 'dodge') + 
  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + 
  # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```

Figure 3. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.

More Americans think there should be much more or or somewhat more fossil fuels used than Danes. More Danes thinks threre should be somewhat less or much less fossil fuels used. This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less. Lets see in pie char form.

```{r,warning=FALSE}
ggplot(fuels, aes(x="", y= Number, fill=response))+  
  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) +
  # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 8. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.

The pie chart shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used. These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%. More Americans believed there should be much more or somewhat more than Danes. Lets verify with a statistical test.

we will run another generalized linear model.

#### Running the Statistical Test

```{r, echo=FALSE}
dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
fuels$num_people<- dumb_way #put that vector into here

fuels<- fuels %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

fuels$total<- round(fuels$total, digits=0) #round to whole number

fuels_numPeople<- fuels %>% select(response, Country, total)
```

Creating a matrix by hand
```{r, echo=FALSE}
t_Den<- c(24,50, 87, 293,590,54,13)
t_US<-c(306,296,445,555,1253,201,39)
t_matrix<- as.matrix(rbind(t_Den, t_US))
chisq.test(t_matrix)
```
Once again, there is a significant difference in how Danes and Americans answered this survey question.

## Summary & Discussion

Key Findings

### Emissions
- Denmark significantly less per capita emissions than U.S.
- Rates of the two countries were not significant

### Survey Questions
- Americans and Danes answered significantly different from eachother in every survey question
- Appears  more Danes think climate change is human caused than Americans
- Appears more Danes want decrease in fossil fuels more than Americans
- Pretty close call regarding which country were more worried or viewed it as a bigger threat

## Limitations
The data used is from a survey conducted on Facebook, so all respondents are Facebook users. This adds a confounding variable, when utilizing this study.

## Challenges

biggest challenge as you have probably guess, is were not gleaning much from the survey data.

want to try ***Diverging stacked bar chart instead***

I am going to try to force it into Likert data, graph diverging bars and do chi squared tests with the negative response grouped in one and the positive grouped in another.

Prior to this project how to conduct a time series analysis, work with survey data, or make pie charts was not discussed nor taught. External research and problem solving had to be conducted in order to achieve this data analysis.

When constructing graphs the pie charts proved to be particularly cumbersome as the percentages were not align on the pie chart in the correct spot, nor all be visible. Research into geom_label and geom_text yeilded a the most succesful outcome. However, the pie charts still need work and any feedback regarding how to would be greatly appreciated.

Research into how to make a likert graph was also conducted. The responses to the four survey questions would have to be the same however, in order to produce a likert graph. I could change the answers to create a made up response scale of pro or anti climate action. Would that be professional or no? Let me know your thoughts so I can include it in the final.

### 2022-11-30 Code/Drafts/LongVersion_Oral_Presentation.Rmd
Oral Presentation Preparation

--- 2022-11-30 Code/Drafts/Final_Short_Oral_Presentation.Rmd
@@
 >CAIT Data by (World Resource Institute) WRI from 1990-2019 including all emissions and per capita
-
->Survey Data 2022 from Yale on Facebook, was in wide format but had more questions
+> Survey Data 2022 from Yale on Facebook, was in wide format but had more questions
 
@@
 
-```{r, echo=FALSE}
+```{r}
 emissions$year<-paste0("01-01-", emissions$year)
@@
 
-Lets determine if there is a difference in rate from when both country's emissions started significantly decreasing and the rate is consistently negative 2010.
-```{r, echo=FALSE}
+Lets determine if there is a difference in rate from when both country's emissions started significantly decreasing and the rate is consistenly negative 2010.
+```{r}
 emissions_rate_2010<- emissions_rate %>% 
@@
 
+Though more people may have been interviewed in one country than the other, the survey data is in percent, as in the percentage out of the whole who chose that option.
+We can double check to ensure the recorded numbers equals 100 per country per survey question.
+
+Lets check with the 'beliefs question' which is the first question we will analyzed.
+
+```{r}
+#Parsing out just the survey answers regarding climate change beliefs
+beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
+```
+
+#### Beliefs of Climate Change
+
+Lets first see what Danes and Americans believe caused climate.
+The options to choose from were: caused mostly by humans, caused by natural change in the environment, caused by both, and climate change is not happening.
+
+```{r, warning=FALSE, echo=FALSE}
+# Making the graph
+ggplot(beliefs, aes(response,Number, fill=Country))+
+  geom_bar(stat = 'identity', position = 'dodge') + 
+ labs(title="Cause of Climate Change", x="Response to Survey Question",
+      y="Percent") + # titling and renaming x and y axis labels
+  coord_flip()+
+  theme_bw()
+
+```
+
+Figure 1. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.
+
+More Americans believe that climate change is not happening or that it is caused by natural changes than Danes.
+More Danes believe it is caused by human activities.
+However, more Danes than Americans think it is cause equally by both.
+Lets look at a pie chart as well.
+
+```{r, warning=FALSE, results='hold'}
+piecolor<- c("orange","maroon","darkgreen","lightblue", "gray70","purple")
+
+beliefs<- beliefs %>% arrange(desc(Number)) %>% 
+  mutate(Position=cumsum(Number/2)-0.5*Number) # moving the percentage labels around
+legend<- c("caused equally","caused mostly by humans",
+           "caused mostly by natural changes","climate change isin't happening",
+           "other","refused") #see below
+ggplot(beliefs, aes(x="", Percent, y= Number, fill=response))+  
+  geom_bar(stat="identity", width=1, color="white")+
+  coord_polar("y", start=0) + 
+  facet_wrap(~Country)+
+ #legend( 3, legend=legend)+, attempt to make the legends smaller, 
+  #so the pie charts could be bigger and I could move the percentages.
+  scale_fill_manual(values=piecolor )+ # manually fill the color
+   theme_void()+# remove background, grid, numeric labels
+  geom_label(aes (label = round(Number, digits = 1), x=1.7),color="white", size=4) 
+# add percentages
+
+```
+
+Figure 2. Pie chart of what Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.
+
+The pie chart shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.
+
+Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.
+
+We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a ***chi-squared test*** after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.
+
+1101 people answered in Denmark
+3096 answered in United States
+
+### Running the Statistical Test
+```{r}
+dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096) 
+#1101 Danes answered the survey, 3096 Americans did.
+beliefs$num_people<- dumb_way #put that vector into here
+
+beliefs<- beliefs %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+beliefs$total<- round(beliefs$total, digits=0) #round to whole number
+
+beliefs_numPeople<- beliefs %>% select(response, Country, total)
+```
+
+Creating a matrix by hand
+```{r}
+vector_Den<- c(585,57,419,19,16,1)
+vector_US<-c(1295,473,1073,86,161,7)
+matrix<- as.matrix(rbind(vector_Den, vector_US))
+chisq.test(matrix)
+```
+There is a significant difference in how Danes and Americans answered this survey question.
+
+### Level of worriedness about Climate Changed
+
+In this survey question people were asked how worried they are about climate change. There are three options for 
+```{r, echo=FALSE}
+worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
+```
+
+```{r, echo=FALSE}
+ggplot(worry, aes(response,Number, fill=Country))+
+  geom_bar(stat = 'identity', position = 'dodge') + 
+  labs(title="Level of Concern Over Climate Change", 
+       x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
+```
+
+Figure 3. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
+
+32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.
+
+```{r, echo=FALSE}
+ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity",
+    width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + 
+  # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+
+Figure 4. Pie chart of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
+
+We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all.
+Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%).
+More Americans however are very worried, and more Americans are not at all worried than Danes.
+Interesting.
+A statistical test is needed to see if Americans are less worried than Danes.
+
+#### Lets do a statistical test to verify.
+
+```{r, echo=FALSE}
+dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096) 
+#1101 Danes answered the survey, 3096 Americans did.
+worry$num_people<- dumb_way #put that vector into here
+
+worry<- worry %>% group_by(Country, response) %>%  
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+worry$total<- round(worry$total, digits=0) #round to whole number
+
+worry_numPeople<- worry %>% select(response, Country, total)
+```
+
+Creating a matrix by hand
+```{r, echo=FALSE}
+w_Den<- c(228,561,261,49,1)
+w_US<-c(991,1102,502,498,3)
+w_matrix<- as.matrix(rbind(w_Den, w_US))
+chisq.test(w_matrix)
+```
+There is a significant difference in how Danes and Americans answered this survey question.
+
+### Is Climate Change a Big Threat
+
+In this survey question people were asked whether or not climate change is a big threat within the next 20 years.
+The options were very serious threat, somewhat serious, not a threat at all, or I do not know.
+
+```{r, echo=FALSE}
+threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
+```
+
+```{r, echo=FALSE}
+ggplot(threat, aes(response,Number, fill=Country))+
+  geom_bar(stat = 'identity', position = 'dodge') + 
+  labs(title="Perceieved Threat of Climate Change", 
+       x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
+```
+
+Figure 5. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
+
+More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans.
+More Americans also think that climate change is not a threat at all than Danes.
+Lets view a pie chart.
+
+```{r, echo=FALSE}
+ggplot(threat, aes(x="", y= Number, fill=response))+  
+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
+   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.5, nudge_y=2) + 
+  # add percentages
+  theme_void()+ # remove background, grid, numeric labels
+facet_wrap(~Country)
+```
+
+Figure 6. Pie char of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
+
+We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat.
+Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that.
+Lets test with a statistical test.
+
+#### Lets do a statistical test to verify.
+
+```{r, echo=FALSE}
+dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096) 
+#1101 Danes answered the survey, 3096 Americans did.
+threat$num_people<- dumb_way #put that vector into here
+
+threat<- worry %>% group_by(Country, response) %>%  mutate(total=(Number/100)*num_people)
+#calculating rate where Number=percentage
+
+threat$total<- round(threat$total, digits=0) #round to whole number
+
+threat_numPeople<- threat %>% select(response, Country, total)
+```
+
+Creating a matrix by hand
+```{r, echo=FALSE}
+t_Den<- c(228,561,261,48,1)
+t_US<-c(991,1102,502,498,3)
+t_matrix<- as.matrix(rbind(t_Den, t_US))
+chisq.test(t_matrix)
+```
+Again, there is a significant difference in how Danes and Americans answered this survey question.
+
 ### More or less Fossil Fuels
 
-This survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.
+Now that we know what Danes and Americans think about climate change and how worried they are, lets see how they are reacting.
+This next survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.
 
-```{r,echo=FALSE}
+Lets check with the 'beilefs question' which is the first we will anaylzed.
+
+```{r, echo=FALSE}
 #Parcing out just the survey answers regarding climate change beliefs
@@
 
-```{r, warning=FALSE, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 # Making the graph
@@
 
-Figure 3. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.
+Figure 7. The fossil fuel usage Danes and Americans think their area should use. Response is in percentage of the total surveyed.
 
@@
 
-```{r,warning=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 ggplot(fuels, aes(x="", y= Number, fill=response))+  
@@
 
-want to try ***Diverging stacked bar chart instead***
+want to try Diverging stacked bar chart instead
 

### 2022-11-30 Code/Drafts/Final_Short_Oral_Presentation.Rmd
changing headers

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
 
-## Introduction
+# Introduction
 
@@
 
-## Preparing R
+# Preparing R
 ```{r, message=FALSE}
@@
 
-## Statistcal Analysis and Results
+# Statistcal Analysis and Results
 
@@
 
-### Climate Change Emissions of Countries
+## Climate Change Emissions of Countries
 
@@
 
-#### Creating the figure
+### Creating the figure
 
@@
 
-#### What about the Rate of Decrease?
+### What about the Rate of Decrease?
 
@@
 
-#### Running the Test
+### Running the Test
 Lets run a paired t test
@@
 
-### Perceptions of Climate Change
+## Perceptions of Climate Change
 
@@
 
-### More or less Fossil Fuels
+## More or less Fossil Fuels
 
@@
 
-#### Running the Statistical Test
+### Running the Statistical Test
 
@@
 
-## Summary & Discussion
+# Summary & Discussion
 
@@
 
-## Limitations
+# Limitations
 The data used is from a survey conducted on Facebook, so all respondents are Facebook users. This adds a confounding variable, when utilizing this study.
 
-## Challenges
+# Challenges
 

### 2022-12-06 Code/Drafts/Final_Short_Oral_Presentation.Rmd
Trying Diverging graphs and new chi squared method

@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.]
-
-### What Data To Use?
-
->CAIT Data by (World Resource Institute) WRI from 1990-2019 including all emissions and per capita
-
->Survey Data 2022 from Yale on Facebook, was in wide format but had more questions
 
@@
 ### 4. If they believe fossil fuel usage should decrease, increase, or stay the same.
+
+### What Data To Use?
+
+>CAIT Data by (World Resource Institute) WRI from 1990-2019 including all emissions and per capita
+
+>Survey Data 2022 from Yale on Facebook, was in wide format but had more questions
 
@@
 
+```{r}
+library(scales)
+fuels<- fuels %>% mutate(percent=paste0(round(Number, digits=1), "%"))
+ggplot(fuels, aes(x=Country, y=Number, fill=response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+  scale_fill_viridis_d()+
+  labs(title="How much Fossil Fuels Should We Use", x=NULL, fill=NULL)
+
+```
+```{r}
+fuels_diverging <- fuels %>% mutate(percent=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number)) 
+#everything in there 
+```
+
+```{r}
+fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(percent, digits=1), "%"))
+ggplot(fuels_diverging, aes(x=Country, y=Number, fill=response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+  scale_fill_viridis_d()+
+  labs(title="How much Fossil Fuels Should We Use", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+```
+
 ```{r,warning=FALSE}
@@
 
+```{r}
+ugh<- survey %>% filter()
+```
+
+```{r}
+fuels_numPeople<-fuels_numPeople %>% filter(response != "Don't know") #excluding don't know
+
+fuels_numPeople$grouped_answer<- NA
+more<- which( fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less" )
+less<- which( fuels_numPeople$response== "Much more" |fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Refused" |fuels_numPeople$response== "Same amount as today")
+
+  fuels_numPeople$grouped_answer[more]<- "more"
+  fuels_numPeople$grouped_answer[less]<- "less"
+
+fuels_numPeople<- fuels_numPeople %>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+fuels_numPeople<- fuels_numPeople %>% filter(grouped_answer != is.na)
+```
+
 # Summary & Discussion

### 2022-12-06 Code/Drafts/Final_Short_Oral_Presentation.Rmd
Transfering Diverging bar and chi squared to Final version

@@
 ```
+
 ```{r}
@@
 
-```{r}
-ugh<- survey %>% filter()
-```
-
-```{r}
-fuels_numPeople<-fuels_numPeople %>% filter(response != "Don't know") #excluding don't know
-
-fuels_numPeople$grouped_answer<- NA
-more<- which( fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less" )
-less<- which( fuels_numPeople$response== "Much more" |fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Refused" |fuels_numPeople$response== "Same amount as today")
-
-  fuels_numPeople$grouped_answer[more]<- "more"
-  fuels_numPeople$grouped_answer[less]<- "less"
-
-fuels_numPeople<- fuels_numPeople %>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
-
-fuels_numPeople<- fuels_numPeople %>% filter(grouped_answer != is.na)
-```
-
 # Summary & Discussion

### 2022-12-13 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Fixing the metadata and the introduction

--- 2022-11-29 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
@@
 
-In the great pursuit of mitigating climate change, countries have differed greatly in their strategies and successes.
+In the great pursuit of mitigating climate change, countries have differed in their strategies and successes.
 Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D.
@@
 U., "What Shapes perceptions of climate change?," WIRES Climate Change.2010] What determines people's perceptions and willingness to decrease their own emissions, is therefore important for countries to understand.
-To delve into this question data from two global surveys on climate perceptions will be analyzed along with green house gas emissions and the reduction rate of countries.
-Special emphasis will be on comparing Denmark due to the countrie's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings.
+To delve into this question I will anaylze data from two global surveys on climate perceptions along with green house gas emissions and the reduction rate of countries.
+Special emphasis will be on comparing Denmark due to the country's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings.
 2022.]
 
-I might use active voice "I will analyze data" vs. "data will be analyzed"
+This data analysis will consist of two sections: the greenhouse gas emissions of the U.S. and Denmark, and the climate change perceptions of their citizens.
 
-This data analysis will consist of three sections: the greenhouse gas emissions of the U.S.
-and Denmark, the climate change perceptions of their citizens, and the correlation between emissions and climate change perceptions, using data from many countries.
-
-In the emissions section Denmark's and the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U.
+In the emissions section Denmark's and the United State's per capita emissions and the rate of decrease will both be compared.I hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U.
 S.
 had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]
-
-Go ahead and use "I" rather than "We" because this is  your work.
 
@@
 These will include what Americans and Danes believe caused climate change, how worried they are about it, how big of a threat they believe climate change is in the next 20 years, and if they believe fossil fuel usage should decrease, increase, or stay the same.
-We hypothesize Denmark will have a significantly higher percentage of people who believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane.
+I hypothesize Denmark will have a significantly higher percentage of people who believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane.
 “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502.
-DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U.
+DOI: 10.22230/cjc.2011v36n3a2367] I further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U.
 S.
-citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, we hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^
-
-In the third section we will overlay the survey data with the emissions data to determine the percentage of causation.
-We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang.
-C., " The influence of political ideology on greenhouse gas emissions" Global Environmental Change 74, no.1 (May 2022):102496.
-DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]
+citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, I hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^
 

### 2022-12-13 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
putting Links in intro

@@
 
-This data analysis will consist of two sections: the greenhouse gas emissions of the U.S. and Denmark, and the climate change perceptions of their citizens.
+This data analysis will consist of two sections: the greenhouse gas emissions of the U.S.and Denmark, and the climate change perceptions of their citizens.
 
@@
 S.
-had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]
+had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]The emissions data data can be found here:  <https://www.climatewatchdata.org/ghg-emissions?end_year=2019&start_year=1990>
 
@@
 S.
-citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, I hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^
+citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not uniformly depicted in this manner in the U.S.^4^ Lastly, I hypothesize that more Danes will answer that fossil fuel usage should be reduced than Americans, due to the aforementioned reasons.^(1,3,4)^ The facebook data used for this section can be found here:<https://data.humdata.org/dataset/climate-change-opinion-survey>.
 

### 2022-12-14 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Finished Final CAP DAP.

@@
 ```
+There is alot of noise, and no discernable trends, so lets look at it smoothed over:
+
+```{r}
+ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
+  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
+#group the points by country so geom_line will draw lines connecting the 
+#data by country.
+
+```
+Denmark's rate is more variable over time, but it stays below 0, after 1995, wherease the US is not consistently below zero until 2000.
+It also goes above zero at 1990.
+Therefore, Denmark started decreasing its emissions five years prior to the U.S.
+and the U.S.'s emissions might be increasing again in 2019.
+Given Denmark has a lower rate from 2010-2019, it may have a significantly lower rate than the U.S.
+Lets run a statistical test to verify.
 
@@
 ```
+Looking at the entire range from 1990-2019, there is no significant difference in rate between Denmark and the U.S. Lets look at 2010-2019 given the differences in the funcitons listed above.
 
-```{r}
-t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
-```
-
-We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series).
-This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.
-
-#### Running the Test
-
-First determine the minimum date
-```{r}
-min_date=min(emissions$year)
-max_date=max(emissions$year)
-```
-I'm not sure you'll need the above code chunk
-
-Turning our data into a time series
-
-```{r}
-emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-```
-
-Lets do time series for the indiviual countries
-```{r}
-emissions_Denmark<- emissions %>% filter(country=="Denmark")
-emissions_US<-emissions %>% filter(country=="United States")
-emissions_Denmark_ts<-ts(emissions_Denmark$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-emissions_US_ts<-ts(emissions_US$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-```
-
-Viewing the time series of each individual country to ensure its correct.
-```{r, results='hide'}
-both_ts<-cbind(emissions_Denmark_ts, emissions_US_ts)
-head(both_ts) # ensuring it worked
-plot(both_ts, xlab="year", ylab="emissions")
-```
-
-If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:
-
-```{r}
-demo_plot<-autoplot(emissions_US_ts)
-demo_plot+
-  labs(x = "Year", y = "Greenhouse gas emissions")+
-  theme_bw()
-```
-
-And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:
-
-```{r}
-both<-cbind(emissions_Denmark_ts, emissions_US_ts)
-autoplot(both)
-```
-
-Here is our data smoothed over so we can see the trends without added noise. The bottom one I presume is denmark. We can see that it is decreasing more.
-
-Making a moving average:
-```{r}
-Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
-Ma_emissions_Denmark<- forecast:: ma(emissions_Denmark_ts, order=3, centre=TRUE)
-Ma_emissions_US<- forecast::ma(emissions_US_ts, order=3, centre=TRUE)
+```{r, echo=FALSE}
+emissions_rate_2010<- emissions_rate %>% 
+  filter(emissions_rate$year>="2010-01-01")
 
@@
 
-Plotting both the moving average and the time series.
-
-```{r, collapse=TRUE}
-plot(emissions_ts)
-lines(Ma_emissions, col="maroon", lwd=3)
-
-plot(emissions_Denmark_ts)
-lines(Ma_emissions_Denmark, col="orange", lwd=3)
-
-plot(emissions_US_ts)
-lines(Ma_emissions_US, col="blue")
-```
-Plotting both with autoplot/ggplot:
-
-```{r}
-USA<-cbind(emissions_US_ts, Ma_emissions_US)
-autoplot(USA)
+```{r, results='hold', echo=FALSE}
+t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
+#complete.cases means just use the ones without NAs
 ```
 
-The red line is the data but smoothed, so we can see the overall trend of the emissions data combined.
-the other two are them separated, notice how similar Denmark's is to the combined data.
-Interesting.
+The p value is smaller at 0.083, however it is still not signifciant. So, we must accept the Null hypothesis. There is no significant difference in rate between Denmark and the United States.
 
-***PUT RUNNING THE TIME SERIES HERE***
-This will really be the paired t-test
+#### Conclusion for Section One
+
+The United States emits significantly more greenhouse gases per capita than Denmark, but there was not significant difference in how fast both countries decreased their emissions.
+
+Now lets look at some examples that may explain why Denmark emits less greenhouse gases per capita.
 
@@
 
-we will run another generalized linear model.
+Making a Diverging Bar Graph
+```{r}
+fuels_diverging <- fuels %>% mutate(percent=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number)) 
+#everything in there 
+```
+
+```{r}
+fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(percent, digits=1), "%"))
+ggplot(fuels_diverging, aes(x=Country, y=Number, fill=response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+  scale_fill_viridis_d()+
+  labs(title="How much Fossil Fuels Should We Use", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+```
+```{r, echo=FALSE}
+dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+fuels$num_people<- dumb_way #put that vector into here
+
+fuels<- fuels %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+fuels$total<- round(fuels$total, digits=0) #round to whole number
+
+fuels_numPeople<- fuels %>% select(response, Country, total)
+```
+
+```{r}
+
+fuels_numPeople<-fuels_numPeople %>% filter(response != "Don’t know") #excluding don't know
+
+fuels_numPeople$grouped_answer<- NA
+more<- which( fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less" )
+less<- which( fuels_numPeople$response== "Much more" |fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Refused" |fuels_numPeople$response== "Same amount as today")
+
+  fuels_numPeople$grouped_answer[more]<- "more"
+  fuels_numPeople$grouped_answer[less]<- "less"
+
+fuels_numPeople<- fuels_numPeople %>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+```
 
@@
 
-***### INSERT STATISTICAL TEST***
-
-## Survey data Impacting Emissions
+```{r}
+fuels.mat<-xtabs(grouped_number~Country, data=fuels_numPeople)
+fuels.mat
+chisq.test(fuels.mat)
+```
+The p value is 2.2e^-16, so we can reject the Null hypothesis. significantly more Americans think we should keep releasing the same amount of fossil fuels or more fossil fuels than Danes.
 
@@
 Let me know your thoughts so I can include it in the final.
+
+## DIVERGING BAR GRAPH SAVING THIS
+
+```{r}
+beliefs_diverging <- beliefs %>% mutate(percent=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
+#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
+```
+
+```{r}
+# This will make our labels are positive numbers.
+
+beliefs_diverging<- beliefs_diverging %>% mutate(percent=paste0(round(percent, digits=1), "%"))
+ggplot(beliefs_diverging, aes(x=Country, y=Number, fill=response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+  scale_fill_viridis_d()+
+  labs(title="What do you Think Caused Climate Change", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+
+```

### 2022-12-14 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
done with intro!
Finished Final CAP DAP.

@@
 
-Good overall, but also include links to each dataset.
-
 ## Preparing R
@@
 
-Statistical analysis will consist of the aforementioned three sections.
+Statistical analysis will consist of the aforementioned two sections.
 
@@
 
-Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.
+Pulling in the corrected Emissions data.
+For the raw data see the raw data folder on github.
+Please also see the metadata txt files per each data set in the Data for the Code folder and the metadata folder.
 
@@
 ```
+Figure 1. Per capita greenhouse gass emissions od Denmark and the U.S.from 1990 to 2019.
 
@@
 
-Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.
+Running a paired t test will let us compare the rate of decrease in emissions of both these countries.
 
-I talked to Robin Lock about this.
-We're going to do a paired t-test where each year has a pair of values (US, Denmark) and the values will be the RATE of change for that year, rather than the GHG emissions themselves.
+### Running the Statistical Test
 
-Here's some info: https://www.statology.org/paired-samples-t-test-r/
+```{r}
+t.test(emissions~country, data=emissions, paired=TRUE)
+#like running a t test, but tell r the data is paired.
+```
+The p-value is very less than 2.2*10^-16, much less than 0.05, so we can reject our Null hypothesis. Denmark does have significantly less GHG emissions per capita than the United States.
+
+## What about the Rate of Decrease?
+
+Calculating rate
 
@@
 denmark_rate<- emissions %>% filter(country=="Denmark")
-den_rate<- diff(denmark_rate$emissions)
+den_rate<- diff(denmark_rate$emissions) # subtracts n by n-1 to get rate.
 denmark_rate$rate[2:30]<-den_rate
@@
 
+Lets look at the rate of Denmark and the United States.
 ```{r}
@@
 ```
+Figure 2. The rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.
 
+There us alot of added noise, making it difficult to see overal trends. Lets view the rate data smoothed over.
+
+```{r}
+ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
+  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
+#group the points by country so geom_line will draw lines connecting the 
+#data by country.
+
+```
+Figure 3. The smoothed over rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.
+
+Denmark's rate is more variable over time, but it stays below 0, after 1995, wherease the US is not consistently below zero until 2000.
+It also goes above zero at 1990.
+Therefore, Denmark started decreasing its emissions five years prior to the U.S.
+and the U.S.'s emissions might be increasing again in 2019.
+Given Denmark has a lower rate from 2010-2019, it may have a significantly lower rate than the U.S.
+Lets run a statistical test to verify.
+
+### Running the Statistical Test
+Lets run a paired t-test to see if this is significant.
 ```{r}
@@
 
+Looking at the entire range from 1990-2019, the p value is 0.7668, so we must accept the Null hypothesiss.
+There is no significant difference in rate between Denmark and the U.S.
+Lets look at 2010-2019 given the differences in the funcitons listed above.
+
+```{r, echo=FALSE}
+# lets just included 210-2019 data.
+emissions_rate_2010<- emissions_rate %>% 
+  filter(emissions_rate$year>="2010-01-01")
+```
+
 ```{r}
@@
 ```
+The p value is smaller at 0.083, however it is still not signifciant. So, we must accept the Null hypothesis. There is no significant difference in rate between Denmark and the United States.
 
-We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series).
-This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.
+### Conclusion for Section One
 
-#### Running the Test
+The United States emits significantly more greenhouse gases per capita than Denmark, but there was no significant difference in how fast both countries decreased their emissions from 1990-2019.
 
-First determine the minimum date
-```{r}
-min_date=min(emissions$year)
-max_date=max(emissions$year)
-```
-I'm not sure you'll need the above code chunk
+Now lets look at some examples that may explain why Denmark emits less greenhouse gases per capita.
 
-Turning our data into a time series
-
-```{r}
-emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-```
-
-Lets do time series for the indiviual countries
-```{r}
-emissions_Denmark<- emissions %>% filter(country=="Denmark")
-emissions_US<-emissions %>% filter(country=="United States")
-emissions_Denmark_ts<-ts(emissions_Denmark$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-emissions_US_ts<-ts(emissions_US$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
-```
-
-Viewing the time series of each individual country to ensure its correct.
-```{r, results='hide'}
-both_ts<-cbind(emissions_Denmark_ts, emissions_US_ts)
-head(both_ts) # ensuring it worked
-plot(both_ts, xlab="year", ylab="emissions")
-```
-
-If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:
-
-```{r}
-demo_plot<-autoplot(emissions_US_ts)
-demo_plot+
-  labs(x = "Year", y = "Greenhouse gas emissions")+
-  theme_bw()
-```
-
-And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:
-
-```{r}
-both<-cbind(emissions_Denmark_ts, emissions_US_ts)
-autoplot(both)
-```
-
-Here is our data smoothed over so we can see the trends without added noise. The bottom one I presume is denmark. We can see that it is decreasing more.
-
-Making a moving average:
-```{r}
-Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
-Ma_emissions_Denmark<- forecast:: ma(emissions_Denmark_ts, order=3, centre=TRUE)
-Ma_emissions_US<- forecast::ma(emissions_US_ts, order=3, centre=TRUE)
-
-```
-
-Plotting both the moving average and the time series.
-
-```{r, collapse=TRUE}
-plot(emissions_ts)
-lines(Ma_emissions, col="maroon", lwd=3)
-
-plot(emissions_Denmark_ts)
-lines(Ma_emissions_Denmark, col="orange", lwd=3)
-
-plot(emissions_US_ts)
-lines(Ma_emissions_US, col="blue")
-```
-Plotting both with autoplot/ggplot:
-
-```{r}
-USA<-cbind(emissions_US_ts, Ma_emissions_US)
-autoplot(USA)
-```
-
-The red line is the data but smoothed, so we can see the overall trend of the emissions data combined.
-the other two are them separated, notice how similar Denmark's is to the combined data.
-Interesting.
-
-***PUT RUNNING THE TIME SERIES HERE***
-This will really be the paired t-test
-
-### Perceptions of Climate Change
+## Perceptions of Climate Change
 
@@
 
-Lets check with the 'beliefs question' which is the first question we will analyzed.
+Lets check with the 'beliefs question' which is the first question we will analyze to ensure Denmark and the U.S. are comparable despite size differences.
 
@@
 beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs")
+
+#Resetting the levels
+
+beliefs$response<-factor(beliefs$response)
+levels(beliefs$response)
 ```
@@
 total
-```
-
-This means we can fairly compare Danes and American's perceptions of climate change.
-
-I talked with Prof Shuckers about these data.
-He thinks we should do the back-calculating to change percents to numbers and then do a chi-square to compare answers between Danes and Americans.
-You will need the data on number of respondants for each question to do that.
-
-???______
-1101 people answered in Denmark
-3096 answered in United States
-
-```{r}
-beliefs <- beliefs %>% group_by(Country)
-
-dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
-beliefs<- cbind(beliefs, dumb_way)
-beliefs<-beliefs %>% rename(num_people=...5)
-
-beliefs<- beliefs %>% group_by(Country, response) %>%  mutate(total=(Number/100)*num_people)
-
-beliefs$total<- round(beliefs$total, digits=0)
-
-test<- beliefs %>% select(response, Country, total)
-```
-
-???? The chi-squared below did not work!!!
-Chi-squared
-```{r}
-beliefs_matrix<-xtabs(total~Country + response, data=test)
-beliefs_matrix
-beliefs_df<- as.data.frame(beliefs_matrix)
-beliefs_df<- beliefs_df %>% filter(Freq>0)
-beliefs_matrix_real<- xtabs(Freq~Country+response, data=beliefs_df)
-beliefs_matrix_real
-chisq.test(beliefs_matrix)
-
-```
-???___
-
-```{r}
-vector_Den<- c(585,57,419,19,16,1)
-vector_US<-c(1295,473,1073,86,161,7)
-matrix<- as.matrix(rbind(vector_Den, vector_US))
-chisq.test(matrix)
-
 ```
@@
 
-Figure 1. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.
+Figure 4. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.
 
@@
 However, more Danes than Americans think it is cause equally by both.
-Lets look at a pie chart as well.
 
-If you want to test these questions with a statistical test, we will need a combination of chi-square tests. Let's talk.
+To get a better sense of the of which country answeres were more beneficial for combating climate change, lets look at a diverging bar graph.
+In this graph answers that increase our chances of combating climate change, in this case believing that climate change is caused by human activity, is made positive and answers that decrease our chances of combating climate change, in this case all other answers, are made negative.
+Due to the small percentage of people who refused, those will be removed from the diverging bar graph, so it is more visually appealing.
 
 ```{r}
-piecolor<- c("orange","maroon","darkgreen","lightblue", "gray70","purple")
+beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
+#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
+```
 
-beliefs<- beliefs %>% arrange(desc(Number)) %>% mutate(Position=cumsum(Number/2)-0.5*Number)
-legend<- c("caused equally","caused mostly by humans","caused mostly by natural changes","climate change isin't happening","other","refused") #see below
-ggplot(beliefs, aes(x="", Percent, y= Number, fill=response))+  
-  geom_bar(stat="identity", width=1, color="white")+
-  coord_polar("y", start=0) + 
-  facet_wrap(~Country)+
- #legend( 3, legend=legend)+, attempt to make the legends smaller, so the pie charts could be bigger and I could move the percentages.
-  scale_fill_manual(values=piecolor )+
-   theme_void()+# remove background, grid, numeric labels
-  geom_label(aes (label = round(Number, digits = 1), x=1.7),color="white", size=4)  # add percentages
+```{r}
+beliefs_diverging_Order<-beliefs_diverging %>% mutate(response=fct_relevel(response, "Caused about equally by human activities and natural changes", "Other", "Caused mostly by natural changes in the environment", "None of the above because climate change isn’t happening", "Caused mostly by human activities"), response=fct_rev(response))
+
+beliefs_diverging_Order
+```
+
+```{r}
+
+#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
+beliefs_diverging_Order<- beliefs_diverging_Order %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+
+beliefs_divering<-ggplot(beliefs_diverging_Order, aes(x=Country, y=Number, fill= response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+
+  #reorganizes the labels and color them
+  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening","Caused mostly by natural changes in the environment", "Other", "Caused about equally by human activities and natural changes","Caused mostly by human activities"), values=c(
+    "None of the above because climate change isn’t happening"="darkorange4",
+    "Caused mostly by natural changes in the environment" = "darkorange3",
+    "Other" = "darkorange", 
+    "Caused about equally by human activities and natural changes"= "goldenrod1",
+    "Caused mostly by human activities" = "deepskyblue"
+  ))+
+
+  labs(title="What do you Think Caused Climate Change", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+
+beliefs_divering
+
+ggsave("beliefs_diverging.png")
+
+```
+Figure 4.
+Diverging Bar graph of what Danes and Americans believe caused climate change.
+Responses are in percent of the total surveyed.Darker orange is a more anti climate change answer, wherease blue is a climate activist answer.
+
+For a easier to read figure please see the image saved in the code folder.
+
+The diverging bar graph shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.
+
+Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.
+
+### Running the Statistical Test
+
+We can put our data into two categories: those that believe climate change is human caused, and all other answers. This way we can run a chi-squared test and see if if it is significant.
+
+first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.
+
+```{r, echo=FALSE}
+respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+beliefs$num_people<- respondants #put that vector into here
+
+beliefs<- beliefs %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+beliefs$total<- round(beliefs$total, digits=0) #round to whole number
+
+beliefs_numPeople<- beliefs%>% select(response, Country, total)
+```
+
+```{r}
+
+beliefs_numPeople$grouped_answer<- NA
+good<- which( beliefs_numPeople$response== "Caused mostly by human activities")
+bad<-which(beliefs_numPeople$response=="Caused mostly by natural changes in the environment" |beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
+           |beliefs_numPeople$response=="Other"
+           |beliefs_numPeople$response=="None of the above because climate change isn’t happening")
+  beliefs_numPeople$grouped_answer[good]<- "good"
+  beliefs_numPeople$grouped_answer[bad]<-"bad"
+
+beliefs_numPeople<- beliefs_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
 
@@
 
-Figure 2. Pie chart of what Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.
+```{r}
+beliefs.mat<-xtabs(grouped_number~Country + grouped_answer, data=beliefs_numPeople)
+beliefs.mat
+chisq.test(beliefs.mat)
+```
+The p values is 3.829*10^-11, which is much less than 95%, so I can reject the Null hypothesis. Significantly more Danes think climate change is caused by humans verses other sources than Americans.
 
-Here's an article that talks more about data viz for likert type data.
-It has good arguments for why to avoid pie charts, and has a building series of charts ending with the author's (and my) favorite.
-Just worth considering:
+## Level of worriedness about Climate Changed
 
-https://nightingaledvs.com/visualizing-likert-scale-data-same-data-displayed-seven-different-ways/
-
-The pie chart shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.
-
-Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.
-
-We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a _________
-
-Chi-square test after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.
-
-***INSERT ### Running the Statistical Test***
-
-#### Level of worriedness about Climate Changed
-In this survey question people were asked how worried they are about climate change. There are three options for 
+In this survey question people were asked how worried they are about climate change. There are three options, very worried, somewhat worried, not very worried, not at all worried, and refused.
 ```{r}
@@
 ```{r}
-ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity') + 
+ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
   labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
@@
 
-Figure 3. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
+Figure 5. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
 
-32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.
+32% of Americans are very worried about climate change, whereas only 21% of Danes are.
+But 52% of Danes are somewhat worried at only 36% are.
+To better visualize this lets look at a diverging bar graph.
+
+Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
+```{r}
+worry_diverging <- worry %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), Number, -Number)) 
+#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
+```
 
 ```{r}
-ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
-   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
-  theme_void()+ # remove background, grid, numeric labels
-facet_wrap(~Country)
+worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response, "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), response=fct_rev(response))
+
 ```
 
-Figure 4. Pie chart of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
+```{r}
+
+#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
+worry_diverging<- worry_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+
+ggplot(worry_diverging, aes(x=Country, y=Number, fill= response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+
+  #reorganizes the labels and color them
+  scale_fill_manual( breaks=c("Not at all worried","Not very worried", "Somewhat worried", "Very worried"),
+      values=c(
+    "Not at all worried"="darkorange4",
+    "Not very worried" = "darkorange3",
+    "Somewhat worried"= "deepskyblue",
+    "Very worried" = "deepskyblue3"
+  ))+
+
+  labs(title="How Much Does Climate Change Worry You", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+
+```
+
+Figure 6. Diverging Bar Graph the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.
 
@@
 More Americans however are very worried, and more Americans are not at all worried than Danes.
-Interesting.
-A statistical test is needed to see if Americans are less worried than Danes.
+Based on the diverging bar graph, there is a slight difference between countries, suggesting Danes are more worried than Americans.
+A statistical test is needed to see if this is a significant difference.
 
-#### Lets do a statistical test to verify. 
-Chi-square test. And think about a new data vis.
+### Lets do a statistical test to verify. 
+Chi-square test.
 
-***### INSERT STATISTICAL TEST***
+Very and somewhat worried will be grouped together and not at all worried and not very worried will be grouped together.
+
+first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.
+
+```{r, echo=FALSE}
+respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+worry$num_people<- respondants #put that vector into here
+
+worry<- worry %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+worry$total<- round(worry$total, digits=0) #round to whole number
+
+worry_numPeople<- worry%>% select(response, Country, total)
+```
+
+```{r}
+
+worry_numPeople<- worry_numPeople %>% filter(response!="Refused")
+worry_numPeople$grouped_answer<- NA
+not<- which(worry_numPeople$response== "Not at all worried"|worry_numPeople$response== "Not very worried")
+worried<- which( worry_numPeople$response== "Very worried" |worry_numPeople$response== "Somewhat worried")
+
+  worry_numPeople$grouped_answer[worried]<- "worried"
+  worry_numPeople$grouped_answer[not]<- "not"
+
+worry_numPeople<- worry_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+```
+
+```{r}
+worry.mat<-xtabs(grouped_number~Country+grouped_answer, data=worry_numPeople)
+worry.mat
+chisq.test(worry.mat)
+```
+the p-value is 0.01258, which is less than 0.05. So, we can reject the Null hypothesis. Danes are more worried about climate change than Americans.
 
@@
 
-Figure 5. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
+Figure 7. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
 
-More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans.
+More Americans than Danes think that climate change is a very serious threat, but more Danes think it is a somewhat serious threat than Americans.
 More Americans also think that climate change is not a threat at all than Danes.
-Lets view a pie chart.
+Lets view a diverging bar graph.
+
+Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
+```{r}
+threat_diverging <- threat %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very serious threat", "Somewhat serious threat"), Number, -Number))
+
+```
+
+Reorganize the bars.
+```{r}
+threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response, "Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))
+
+```
 
 ```{r}
-ggplot(threat, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
-   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.5, nudge_y = 2 ) + # add percentages
-  theme_void()+ # remove background, grid, numeric labels
-facet_wrap(~Country)
+#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
+threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+
+ggplot(threat_diverging, aes(x=Country, y=Number, fill= response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+
+  #reorganizes the labels and color them
+  scale_fill_manual( breaks=c("Not a threat at all","Don't know", "Somewhat serious threat", "Very serious threat"),
+      values=c(
+    "Not a threat at all"= "darkorange3",
+    "Don't know" = "goldenrod3",
+    "Somewhat serious threat"= "deepskyblue",
+    "Very serious threat" = "deepskyblue3"
+  ))+
+
+  labs(title="In the Next 20 Years How Much of a Threat is Climate Change?", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+
 ```
 
-Figure 6. Pie char of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
+Figure 8. Diverging bar graph of the perceived threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.
 
-We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat.
+We can see that most Danes, 53%, think climate change is a somewhat serious threat, wherease most Americans, 40.6%, think it is a very serious threat.
 Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that.
-Lets test with a statistical test.
+There is a small shift in Dane's bars leaning right, suggesting they think climate change is a more serious threat than Americans.
+Lets see if it is significant with a statistical test.
 
-#### Lets do a statistical test to verify.
+### Lets do a statistical test to verify.
 
-***###INSERT STATISTICAL TEST ***
+first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.
 
-### More or less Fossil Fuels
+```{r, echo=FALSE}
+respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+threat$num_people<- respondants #put that vector into here
+
+threat<- threat %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+threat$total<- round(threat$total, digits=0) #round to whole number
+
+threat_numPeople<- threat%>% select(response, Country, total)
+```
+
+```{r}
+
+threat_numPeople<- threat_numPeople %>% filter(response!="Refused") 
+threat_numPeople$grouped_answer<- NA
+no<- which(threat_numPeople$response== "Not a threat at all" |threat_numPeople$response== "Don't know" )
+yes<- which(threat_numPeople$response== "Somewhat serious threat" |threat_numPeople$response== "Very serious threat")
+
+  threat_numPeople$grouped_answer[no]<- "no"
+  threat_numPeople$grouped_answer[yes]<- "yes"
+
+threat_numPeople<- threat_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+```
+
+```{r}
+threat.mat<-xtabs(grouped_number~Country+grouped_answer, data=threat_numPeople)
+threat.mat
+chisq.test(threat.mat)
+```
+Pretending 'Don't know' is a negative response, there is not a significant difference in how much of a threat climate change is perceived as by Danes and Americans.
+
+```{r, echo=FALSE}
+respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+threat$num_people<- respondants #put that vector into here
+
+threat2<- threat %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+threat2$total<- round(threat2$total, digits=0) #round to whole number
+
+threat2_numPeople<- threat2%>% select(response, Country, total)
+```
+
+```{r, echo=FALSE}
+
+threat2_numPeople<- threat2_numPeople %>% filter(response!="Refused" & response!="Don't know") 
+threat2_numPeople$grouped_answer<- NA
+no<- which(threat2_numPeople$response== "Not a threat at all" )
+yes<- which(threat2_numPeople$response== "Somewhat serious threat" |threat2_numPeople$response== "Very serious threat")
+
+  threat2_numPeople$grouped_answer[no]<- "no"
+  threat2_numPeople$grouped_answer[yes]<- "yes"
+
+threat2_numPeople<- threat2_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+```
+
+```{r, echo=FALSE}
+threat2.mat<-xtabs(grouped_number~Country+grouped_answer, data=threat2_numPeople)
+threat2.mat
+chisq.test(threat2.mat)
+```
+Ignoring the Don't knows yields a p value of 0.005, so I can reject the Null hypothesis. Significantly more Danes think that within the next 20 years climate change is a serious threat.
+
+## More or less Fossil Fuels
 
@@
 This next survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.
-
-Lets check with the 'beilefs question' which is the first we will anaylzed.
 
@@
 This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less.
-Lets see in pie char form.
+Lets see in diverging bar graph.
+
+Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
+
+Again the option "Don't know" is present.
+This also has the option of same amount as today.
+According to the IPCC humans must decrease their emissions to stay well below 2 degrees of warming and avoid the most severe impacts of climate change, therefore, keeping it the same will go on the negative side of the diverging bar graph.
+Not many people chose "Don't know," plus it is difficult to predict, so I believe it is fair to leave it out.
+```{r}
+fuels_diverging <- fuels %>% filter(response!="Refused" & response!= "Don’t know") %>% mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number))
+
+```
+
+Reorganize the bars.
+```{r}
+fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,"Much more","Somewhat more", "Same amount as today", "Somewhat less", "Much less"), response=fct_rev(response))
+
+```
 
 ```{r}
-ggplot(fuels, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
-   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
-  theme_void()+ # remove background, grid, numeric labels
-facet_wrap(~Country)
+#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
+fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number), digits=0), "%"))
+#rounded it to 0 this time due to serious overlap in the % labels on the graph
+
+ggplot(fuels_diverging, aes(x=Country, y=Number, fill= response)) + 
+  geom_col()+ #also makes a bar graph
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  coord_flip()+
+  scale_x_discrete()+
+
+  #reorganizes the labels and color them
+  scale_fill_manual( breaks=c("Much more", "Somewhat more", "Same amount as today", "Somewhat less", "Much less"),
+      values=c(
+    "Much more"= "darkorange4",
+    "Somewhat more"= "darkorange3",
+    "Same amount as today" = "darkorange2",
+    "Somewhat less"= "deepskyblue",
+    "Much less" = "deepskyblue3"
+  ))+
+
+  labs(title="How Much Fossil Fuels Should Your Country Use?", x=NULL, fill=NULL)+
+ theme_minimal()+
+  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
+panel.grid=element_blank(),
+legend.position="top")
+
 ```
@@
 
-The pie chart shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used.
-These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%.
-More Americans believed there should be much more or somewhat more than Danes.
+The diverging bar graph shows that most of the Danes surveyed, 54%, believed much less fossil fuels should be used, with the second largest percentage of 26% believed somewhat less should be used.
+These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40%, with the second most thinking somewhat less fossil fuel usage, 18%.
+More Americans believed there should be much more or somewhat more fossil fuel usage than Danes.
+There is also a visible difference in the overall diverging bar graphs, sugesting that more Danes believe their country should use less fossil fuels, than Americans believed.
 Lets verify with a statistical test.
 
-we will run another generalized linear model.
+we will run chi-squared test.
 
-#### Running the Statistical Test
+### Running the Statistical Test
 
-***### INSERT STATISTICAL TEST***
+```{r, echo=FALSE}
+respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
+#1101 Danes answered the survey, 3096 Americans did.
+fuels$num_people<- respondants #put that vector into here
 
-## Survey data Impacting Emissions
+fuels<- fuels %>% group_by(Country, response) %>% 
+  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+
+fuels$total<- round(fuels$total, digits=0) #round to whole number
+
+fuels_numPeople<- fuels%>% select(response, Country, total)
+```
+
+```{r, echo=FALSE}
+
+fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & response!="Don't know") 
+fuels_numPeople$grouped_answer<- NA
+more<- which(fuels_numPeople$response== "Much more" | fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Same amount as today" )
+less<- which(fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less")
+
+  fuels_numPeople$grouped_answer[more]<- "more"
+  fuels_numPeople$grouped_answer[less]<- "less"
+
+fuels_numPeople<- fuels_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))
+
+```
+
+```{r, echo=FALSE}
+fuels.mat<-xtabs(grouped_number~Country+grouped_answer, data=fuels_numPeople)
+fuels.mat
+chisq.test(fuels.mat)
+```
+The p-value is 2.2*10^-16, much less than 0.05, meaning we can reject the null hypothesis. Significantly more Danes believe their country should use less fossil fuels, than Americans.
 
@@
 
+Denmark emits significantly less per capita greenhouse emissions than the United states.
+My results also showed that significantly more Danes believed climate change was human caused, were more worried aobut climate change, and think their country should decrease fossil fuel usage than Americans did.
+Overall, Danes' answers to the survey questions were significantly more planet friendly than Americans.
+Danish citizen's perceptions of climate change could be way Denmark emmits so much less greenhouse gas emissions per person than the United States.
+
+It was difficult to asses the to what degree in the next 20 years is climate change a threat question, due to the "don't know" option.
+With it, there was no significant difference in responses between the countries.
+If I do not include it Denmark thought climate change was a bigger threat than the United States.
+On one hand people who answered do not know probably don't partake in climate change related activities such as voting for policies, making their decisions neutral.
+On the other hand, it could mean the person would lean against climate policies, not really considering climate change a pertinent issue to them personally.
+It is difficult to determine how someone who answered don't know would react, so not including it seems reasonable.
+Furthermore, given Denmark answered significantly more favorably in the three other questions, a significant p value when removing the "don't know" responses is not surprising.
+However, it is interesting that more Danes answered did not know to this question than Americans.
+
 ## Limitations
-The data used is from a survey conducted on Facebook, so all respondents are Facebook users. This adds a confounding variable, when utilizing this study.
+
+There were not many limitaitons with the emissions data, aside from missing 2019-2023 emissions, which may have showed an interesting emissions and rate of emisisons correlation.
+
+The data used is from the survey data was conducted on Facebook, so all respondents are Facebook users.
+This adds a confounding variable, when utilizing this study.
+However, the data was weighted to account for any biases, so the limitation does not render the data moot.
+
+It is surprising, however, that the survey data did not have the option  'minor threat' or 'small threat' for the threat within the next 20 years question, to balance out the likert data scale in place of the don't know option.
+The don't know options makes it difficult to properly compare Danes and Americans opinions regarding regarding that question.
 
@@
 
-Prior to this project how to conduct a time series analysis, work with survey data, or make pie charts was not discussed nor taught.
-External research and problem solving had to be conducted in order to achieve this data analysis.
+The two data sets were both in wide format, with the survey data being particularly cumbersome, given each question was on a separate sheet.
+The emissions data also had x's infront of all its date entries and the dates were listed as factor data.
+I used sapply and lapply to fix this, which I had to research and teach myself.
+Overall however, the exploratory anayslis was not as challenging as the next step.
 
-What about challenges in terms of figuring out how to reshape/reformat your data? Anything worth mentioning there?
+Prior to this project how to conduct a two paired t-test, work with survey data, or make diverging bar graphs was not discussed nor taught.
+External research and problem solving had to be conducted in order to achieve this for the final analysis.
 
-This is very thorough and a solid first draft.
-
-When constructing graphs the pie charts proved to be particularly cumbersome as the percentages were not align on the pie chart in the correct spot, nor all be visible.
-Research into geom_label and geom_text yeilded a the most succesful outcome.
-However, the pie charts still need work and any feedback regarding how to would be greatly appreciated.
-
-Research into how to make a likert graph was also conducted.
-The responses to the four survey questions would have to be the same however, in order to produce a likert graph.
-I could change the answers to create a made up response scale of pro or anti climate action.
-Would that be professional or no?
-Let me know your thoughts so I can include it in the final.
+When constructing graphs the diverging bar graphs  proved to be particularly cumbersome as the labels, bars, and bar colors all needed to be in the same appropiate order.
+The first beliefs survey data was difficult for this, because the answers to the survey were so long, they could not all fit on the bar graph.
+Given diverging bar graphs are parced out based on the available factor levels of the data I could not change the long data responses, so instead I provided an image people can view along with the pdf to see the graph better.

### 2022-12-14 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Cleaning up the Markdown

@@
 ```{r}
-emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
+emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", 
+              "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
 ```
@@
 emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
-#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
+#format tells R the dates are just a four digit year, a two digit year 
+#is %y, origin informs R where it starts.
 # must do the - in the %d- because the dates are entered with a -. 
@@
 ```{r}
-ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  geom_line() +
-  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
+ggplot(emissions, aes(year, emissions, colour= country, group=country)) +  
+  geom_line() +
+  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, 
+#and group tells R to group the points by country so geom_line will draw
+#lines connecting the data by country.
 
@@
 
-```{r}
+```{r, results='hold'}
 t.test(emissions~country, data=emissions, paired=TRUE)
@@
 ```{r}
-ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
-  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
+ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) + 
+  geom_line() +
+  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, 
+#and group tells R to group the points by country so geom_line will draw 
+#lines connecting the data by country.
 ```
@@
 
-```{r}
-ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
-  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
+```{r, message=FALSE}
+#colour colors by country, and group tells R to 
 #group the points by country so geom_line will draw lines connecting the 
 #data by country.
+ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
+  geom_smooth() + theme_bw()
 
@@
 ```{r}
+#complete.cases means just use the ones without NAs
 t.test(rate~country, data=emissions_rate, paired=TRUE)
-#complete.cases means just use the ones without NAs
 ```
@@
 There is no significant difference in rate between Denmark and the U.S.
-Lets look at 2010-2019 given the differences in the funcitons listed above.
+Lets look at 2010-2019 given the differences in the functions listed above.
 
@@
 ```{r}
-survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
+survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", 
+                       "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
 ```
@@
 #Resetting the levels
-
 beliefs$response<-factor(beliefs$response)
@@
 
-```{r}
+```{r, warning=FALSE}
 # Making the graph
@@
   geom_bar(stat = 'identity', position = 'dodge') + 
- labs(title="Cause of Climate Change", x="Response to Survey Question", y="Percent") + # titling and renaming x and y axis labels
-  #theme( axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))+ this won't work :(
+ labs(title="Cause of Climate Change", x="Response to Survey Question", 
+      y="Percent") + # titling and renaming x and y axis labels
   coord_flip()+
@@
 ```{r}
-beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
-#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
+beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% 
+  mutate(Number=if_else(response %in% 
+                    c("Caused mostly by human activities"), Number, -Number)) 
+#all the percentages in the response column that did not 
+#answer "caused mostly by human activities"   will have a - put in front. 
 ```
@@
 ```{r}
-beliefs_diverging_Order<-beliefs_diverging %>% mutate(response=fct_relevel(response, "Caused about equally by human activities and natural changes", "Other", "Caused mostly by natural changes in the environment", "None of the above because climate change isn’t happening", "Caused mostly by human activities"), response=fct_rev(response))
-
-beliefs_diverging_Order
+beliefs_diverging_Order<-beliefs_diverging %>% 
+  mutate(response=fct_relevel(response, 
+  "Caused about equally by human activities and natural changes", 
+  "Other", "Caused mostly by natural changes in the environment", 
+  "None of the above because climate change isn’t happening", 
+  "Caused mostly by human activities"), response=fct_rev(response)) 
 ```
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 
-#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
-beliefs_diverging_Order<- beliefs_diverging_Order %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+#first line below takes the Number column rounds it up, 
+#takes the absolute value of it and adds a % for the labels on the graph.
+beliefs_diverging_Order<- beliefs_diverging_Order %>%
+  mutate(percent=paste0(round(abs(Number), digits=0), "%"))
 
-beliefs_divering<-ggplot(beliefs_diverging_Order, aes(x=Country, y=Number, fill= response)) + 
+beliefs_divering<-ggplot(beliefs_diverging_Order,
+                         aes(x=Country, y=Number, fill= response)) + 
   geom_col()+ #also makes a bar graph
-  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5),
+            color="white", fontface="bold")+ 
+  #stack stacks them, fontface bolds  the numbers in the bargraphs 
   coord_flip()+
@@
   #reorganizes the labels and color them
-  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening","Caused mostly by natural changes in the environment", "Other", "Caused about equally by human activities and natural changes","Caused mostly by human activities"), values=c(
+  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening",
+  "Caused mostly by natural changes in the environment", "Other", 
+  "Caused about equally by human activities and natural changes",
+  "Caused mostly by human activities"), 
+  values=c(
     "None of the above because climate change isn’t happening"="darkorange4",
@@
 
+Given  the diverging bar graph and chi squared test code chunks are virtually the same as above, they are not shown in the pdf version henceforth. 
 ```{r, echo=FALSE}
@@
 
-```{r}
+```{r, echo=FALSE}
 
@@
 good<- which( beliefs_numPeople$response== "Caused mostly by human activities")
-bad<-which(beliefs_numPeople$response=="Caused mostly by natural changes in the environment" |beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
+bad<-which(beliefs_numPeople$response==
+             "Caused mostly by natural changes in the environment"
+|beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
            |beliefs_numPeople$response=="Other"
@@
 
-```{r}
+```{r, echo=FALSE}
 beliefs.mat<-xtabs(grouped_number~Country + grouped_answer, data=beliefs_numPeople)
@@
 
-```{r}
-ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
-  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
+```{r, echo=FALSE}
+ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
+                                                          position = 'dodge') + 
+  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", 
+       y="Percent") + 
+  theme_bw() +
+  coord_flip()
 ```
@@
 Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
-```{r}
-worry_diverging <- worry %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), Number, -Number)) 
-#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
-```
-
-```{r}
-worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response, "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), response=fct_rev(response))
+```{r, echo=FALSE}
+worry_diverging <- worry %>% filter(response!="Refused") %>% 
+  mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), 
+                        Number, -Number))
 
@@
 
-```{r}
+```{r, echo=FALSE}
+worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response,
+  "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), 
+  response=fct_rev(response))
+
+```
+
+```{r, echo=FALSE}
 
 #first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
-worry_diverging<- worry_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+worry_diverging<- worry_diverging %>% mutate(percent=paste0(round(abs(Number),
+                                                          digits=1), "%"))
 
@@
   geom_col()+ #also makes a bar graph
-  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
+  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), 
+            color="white", fontface="bold")+ 
+  #stack stacks them, fontface bolds  the numbers in the bargraphs 
   coord_flip()+
@@
   #reorganizes the labels and color them
-  scale_fill_manual( breaks=c("Not at all worried","Not very worried", "Somewhat worried", "Very worried"),
+  scale_fill_manual( breaks=c("Not at all worried","Not very worried", 
+                            "Somewhat worried", "Very worried"),
       values=c(
@@
 worry_numPeople$grouped_answer<- NA
-not<- which(worry_numPeople$response== "Not at all worried"|worry_numPeople$response== "Not very worried")
-worried<- which( worry_numPeople$response== "Very worried" |worry_numPeople$response== "Somewhat worried")
+not<- which(worry_numPeople$response== "Not at all worried"
+            |worry_numPeople$response== "Not very worried")
+worried<- which( worry_numPeople$response== "Very worried" 
+                 |worry_numPeople$response== "Somewhat worried")
 
@@
 
-```{r}
+```{r, echo=FALSE}
 worry.mat<-xtabs(grouped_number~Country+grouped_answer, data=worry_numPeople)
@@
 
-```{r}
-ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
-  labs(title="Perceieved Threat of Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
+```{r, echo=FALSE}
+ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
+                                                        position = 'dodge') + 
+  labs(title="Perceieved Threat of Climate Change", 
+       x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
 ```
@@
 Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
-```{r}
-threat_diverging <- threat %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very serious threat", "Somewhat serious threat"), Number, -Number))
+```{r, echo=FALSE}
+threat_diverging <- threat %>% filter(response!="Refused") %>% 
+  mutate(Number=if_else(response %in% c("Very serious threat", 
+                                "Somewhat serious threat"), Number, -Number))
 
@@
 
-Reorganize the bars.
-```{r}
-threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response, "Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))
+```{r, echo=FALSE}
+threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response,
+"Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))
 
@@
 
-```{r}
-#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
-threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))
+```{r, echo=FALSE}
+#first line below takes the Number column rounds it up, takes the absolute 
+#value of it and adds a % for the labels on the graph.
+threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), 
+                                                              digits=1), "%"))
 
@@
   #reorganizes the labels and color them
-  scale_fill_manual( breaks=c("Not a threat at all","Don't know", "Somewhat serious threat", "Very serious threat"),
+  scale_fill_manual( breaks=c("Not a threat at all","Don't know",
+                            "Somewhat serious threat", "Very serious threat"),
       values=c(
@@
 
-  labs(title="In the Next 20 Years How Much of a Threat is Climate Change?", x=NULL, fill=NULL)+
+  labs(title="In the Next 20 Years How Much of a Threat is Climate Change?", 
+       x=NULL, fill=NULL)+
  theme_minimal()+
@@
 threat<- threat %>% group_by(Country, response) %>% 
-  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage
+  mutate(total=(Number/100)*num_people) 
+#calculating rate where Number=percentage
 
@@
 threat_numPeople$grouped_answer<- NA
-no<- which(threat_numPeople$response== "Not a threat at all" |threat_numPeople$response== "Don't know" )
-yes<- which(threat_numPeople$response== "Somewhat serious threat" |threat_numPeople$response== "Very serious threat")
+no<- which(threat_numPeople$response== "Not a threat at all" 
+           |threat_numPeople$response== "Don't know" )
+yes<- which(threat_numPeople$response== "Somewhat serious threat" 
+            |threat_numPeople$response== "Very serious threat")
 
@@
 
-threat2_numPeople<- threat2_numPeople %>% filter(response!="Refused" & response!="Don't know") 
+threat2_numPeople<- threat2_numPeople %>% filter(response!="Refused" &
+                                                   response!="Don't know") 
 threat2_numPeople$grouped_answer<- NA
 no<- which(threat2_numPeople$response== "Not a threat at all" )
-yes<- which(threat2_numPeople$response== "Somewhat serious threat" |threat2_numPeople$response== "Very serious threat")
+yes<- which(threat2_numPeople$response== "Somewhat serious threat" 
+            |threat2_numPeople$response== "Very serious threat")
 
@@
 
-```{r}
+```{r, warning=FALSE, echo=FALSE}
 # Making the graph
-ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
-  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + # renaming x and y axis labels
+ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity',
+                                                           position = 'dodge') + 
+  labs(title="Fossil Fuel Use", x="Response to Survey Question", y="Percent") + 
+  # renaming x and y axis labels
   theme_bw() +
@@
 Not many people chose "Don't know," plus it is difficult to predict, so I believe it is fair to leave it out.
-```{r}
-fuels_diverging <- fuels %>% filter(response!="Refused" & response!= "Don’t know") %>% mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number))
+```{r, echo=FALSE}
+fuels_diverging <- fuels %>% filter(response!="Refused" 
+                                    & response!= "Don’t know")%>% 
+  mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), 
+                          Number, -Number)) 
+```
+
+```{r, echo=FALSE}
+fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,
+"Much more","Somewhat more", "Same amount as today", "Somewhat less", 
+"Much less"), response=fct_rev(response))
 
@@
 
-Reorganize the bars.
-```{r}
-fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,"Much more","Somewhat more", "Same amount as today", "Somewhat less", "Much less"), response=fct_rev(response))
-
-```
-
-```{r}
-#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
-fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number), digits=0), "%"))
+```{r,echo=FALSE}
+#first line below takes the Number column rounds it up, takes the absolute 
+#value of it and adds a % for the labels on the graph.
+fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number),
+                                                            digits=0), "%"))
 #rounded it to 0 this time due to serious overlap in the % labels on the graph
@@
   #reorganizes the labels and color them
-  scale_fill_manual( breaks=c("Much more", "Somewhat more", "Same amount as today", "Somewhat less", "Much less"),
+  scale_fill_manual( breaks=c("Much more", "Somewhat more", 
+                      "Same amount as today", "Somewhat less", "Much less"),
       values=c(
@@
 
-fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & response!="Don't know") 
+fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & 
+                                               response!="Don't know") 
 fuels_numPeople$grouped_answer<- NA
-more<- which(fuels_numPeople$response== "Much more" | fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Same amount as today" )
-less<- which(fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less")
+more<- which(fuels_numPeople$response== "Much more" |
+  fuels_numPeople$response== "Somewhat more" | 
+    fuels_numPeople$response== "Same amount as today" )
+less<- which(fuels_numPeople$response== "Much less" |
+               fuels_numPeople$response== "Somewhat less")
 

### 2022-12-30 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
tightening it up

@@
 We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all.
-Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%).
+Most Americans are also somewhat worried, though the percentage is smaller compared to Denmark (51% to 35.6%).
 More Americans however are very worried, and more Americans are not at all worried than Danes.
@@
 ```
-Ignoring the Don't knows yields a p value of 0.005, so I can reject the Null hypothesis. Significantly more Danes think that within the next 20 years climate change is a serious threat.
+Removing the 'Don't know'responses  yields a p value of 0.005, so I can reject the Null hypothesis.
+Significantly more Danes think that within the next 20 years climate change is a serious threat.
+See the discussion for more information.
 
@@
 Denmark emits significantly less per capita greenhouse emissions than the United states.
-My results also showed that significantly more Danes believed climate change was human caused, were more worried aobut climate change, and think their country should decrease fossil fuel usage than Americans did.
+My results also showed that significantly more Danes believed climate change was human caused, were more worried abut climate change, and think their country should decrease fossil fuel usage than Americans did.
 Overall, Danes' answers to the survey questions were significantly more planet friendly than Americans.
-Danish citizen's perceptions of climate change could be way Denmark emmits so much less greenhouse gas emissions per person than the United States.
+Danish citizen's perceptions of climate change could be why Denmark emits significantly less greenhouse gas emissions per person than the United States.
 
-It was difficult to asses the to what degree in the next 20 years is climate change a threat question, due to the "don't know" option.
+It was difficult to asses the 'to what degree in the next 20 years is climate change a threat question', due to the "don't know" option.
 With it, there was no significant difference in responses between the countries.
